### 2012-10-06 docs/tex/draft.tex
Skeleton documentation. Also a bash script for compiling the latex file

\documentclass{article}
\usepackage{graphicx}
\usepackage{amsmath,amssymb,amsthm}
%\usepackage[square]{natbib}

\begin{document}

\title{Pareto Optimal Sequence Design using Tree Graphical Models}
\author{Subhodeep Moitra, CMU}

\maketitle

\begin{abstract}
To be added shortly
\end{abstract}

\section{Introduction}
To be added shortly \cite{Ambroggio2006}.

\section{Method}

\begin{equation}
    \label{simple_equation}
    \alpha = \sqrt{ \beta }
\end{equation}

\subsection{Subsection Heading Here}
Write your subsection text here.

%\begin{figure}
%    \centering
%    \includegraphics[width=3.0in]{myfigure}
%    \caption{Simulation Results}
%    \label{simulationfigure}
%\end{figure}

\section{Conclusion}
Write your conclusion here.

\bibliographystyle{plain}
\bibliography{refs}

\end{document}

### 2012-10-06 docs/tex/compile_tex.sh

@@ -1,0 +1,20 @@
+#!/bin/sh
+
+# Bash script to compile latex file
+# Input - name of main tex file
+
+fn=$1
+fbase=`basename ${1} .tex `
+
+echo ".tex filename is $fn"
+echo ".tex basename is $fbase"
+
+pdflatex ${fn}
+bibtex ${fbase}.aux
+pdflatex ${fn}
+pdflatex ${fn}
+
+
+
+
+

### 2012-10-06 docs/tex/draft.tex
Added HMM section

@@
 
+\DeclareMathOperator*{\argmax}{arg\,max}
+
 \begin{document}
 
-\title{Pareto Optimal Sequence Design using Tree Graphical Models}
+\title{Pareto-Optimal Multistate Protein Design using Tree Graphical Models}
 \author{Subhodeep Moitra, CMU}
@@
 \section{Method}
+In this section, we describe and outline aspects of the method.
+The key idea is that being able to find the minimum energy sequence using dynamic programming(DP) allows us to find pareto optimal solution sequences for all convex combinations of energy functions of multi-states.
+Thus, we want to cleverly map sequence to an energy value using a function that can be solved using DP.
+
+\subsection{Hidden Markov Model(HMM)}
+An example of an energy function that can be solved using DP is the popular Hidden Markov Model(HMM).
+The HMM is probabilistic model represented as a chain of hidden markov states where each state has an associated observed variable.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.8\textwidth]{pics/hmm}
+    \caption{A canonical HMM}
+    \label{hmm1}
+\end{figure}
+
+The graphical representation of the HMM factorizes according to the following probability distribution :
+
+\begin{equation}
+P(X_1,\ldots,X_n|\mathbf{O_1},\ldots,\mathbf{O_n}) = P(X_1)\prod_{i=1}^{n}{P(\mathbf{O_i}|X_i)}\prod_{i=1}^{n-1}{P(X_{i+1}|X_i)}
+\end{equation}
+
+Note, that the chain structure gives rise to a distribution where the dependence between random variables $X_1 to X_n$ is only through adjacent random variables $X_i and X_{i+1}$.
+This interesting decomposition allows us to find the minimum energy/maximum probability sequence efficiently using DP.
+The DP solution to a HMM is the famous Viterbi Decoding algorithm.
+In particular, we want the solution of :
+
+\begin{align*}
+X^\star_{1:n} &= \argmax_{X_1,\ldots ,X_n} P(X_1,\ldots,X_n|\mathbf{O_1},\ldots,\mathbf{O_n})\\
+  &= \argmax_{X_1,\ldots ,X_n}P(X_1)\prod_{i=1}^{n}{P(\mathbf{O_i}|X_i)}\prod_{i=1}^{n-1}{P(X_{i+1}|X_i)}
+\end{align*}
+
+The main steps in the Viterbi algorithm are : 
+\\
+\\
+\textit{Calculation of the Optimal Probability value:}
+\[
+  V(i,k) = \left\{
+  \begin{array}{l l}
+    P(\mathbf{O_1}|X_1{=}k)P(X_1{=}k) & \quad \text{if $i=1$}\\
+    P(\mathbf{O_i}|X_i{=}k)\max_{j}\{P(X_{i}{=}k|X_{i-1}{=}j)V(i-1,j)\} & \quad \text{if $2\leq i\leq n$}\\
+  \end{array} \right.
+\]
+\\
+\textit{Retrieval of the optimal sequence corresponding to the optimal probability value:}
+\begin{align*}
+X^\star_n &= \argmax_{k}V(n,k) \\
+X^\star_{i} &= Ptr(X^\star_{i+1},i+1) & \quad \text{if $1\leq i \leq n-1 $}
+\end{align*}
+
+Here, $V(i,k)$ corresponds to the maximum probability value attainable by examining the sequence from $X_{1:i}$ and the observations from $\mathbf{O_{1:i}}$.
+Also, $X^\star_{1:n}$ corresponds to the optimal sequence assignment.
+Pointers are stored to each of the previous variables during the calculation of $V(i,k)$.
+These pointers are used to retrieve the optimal sequence assignment.
+
+\subsection{Chain Structured Markov Random Field (CMRF)}
+We will consider a generalization of the HMM, to an undirected graphical model.
+This is known as the chain structured MRF.
+The purpose of generalizing to an MRF is that it gives more flexibility in the form of the energy functions and also that there are more general inference algorithms that can be used.
+In particular,
+
+\subsection{Convex combinations of Energy Functions}
 
@@
 \end{equation}
-
-\subsection{Subsection Heading Here}
-Write your subsection text here.
 

### 2012-10-06 docs/tex/draft.tex
Added Latex Picture for CMRF

@@
 %\usepackage[square]{natbib}
+\usepackage{color}
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{pics/hmm}
+    \includegraphics[width=0.8\textwidth]{pics/hmm.pdf}
     \caption{A canonical HMM}
@@
 The main steps in the Viterbi algorithm are :
-\\
-\\
+
+\vspace*{0.2in}
+
 \textit{Calculation of the Optimal Probability value:}
@@
 \]
-\\
+
 \textit{Retrieval of the optimal sequence corresponding to the optimal probability value:}
@@
 We will consider a generalization of the HMM, to an undirected graphical model.
-This is known as the chain structured MRF.
+This is known as the chain structured Markov Random Field (CMRF).
 The purpose of generalizing to an MRF is that it gives more flexibility in the form of the energy functions and also that there are more general inference algorithms that can be used.
-In particular,
+In particular, the Max-Product algorithm can be used to solve the CMRF.
+The CMRF replaces the Conditional probability tables(CPTs) present in the HMM with node and edge potentials.
+The following figure describes the CMRF.
+
+\begin{figure}[h]
+% selecting unit length
+\centering
+% used for centering Figure
+\setlength{\unitlength}{0.3cm}
+\begin{picture}(40,20)(0,0)
+\multiput(5,15)(8,0){2}{\circle{4}}
+\multiput(5,7)(8,0){2}{\circle{4}}
+\multiput(5,7)(8,0){2}{\circle{3.7}}
+\thicklines
+\multiput(7,15)(8,0){2}{\line(1,0){4}}
+\multiput(20,15)(1,0){4}{\circle*{0.3}}
+\multiput(5,13)(8,0){2}{\line(0,-1){4}}
+\put(24,15){\line(1,0){4}}
+\put(30,13){\line(0,-1){4}}
+\thinlines
+\put(30,15){\circle{4}}
+\put(30,7){\circle{4}}
+\put(30,7){\circle{3.7}}
+\put(3.5,14.6){$\phi(X_1)$}
+\put(11.5,14.6){$\phi(X_2)$}
+\put(28.5,14.6){$\phi(X_n)$}
+\put(3.5,6.6){$\phi(O_1)$}
+\put(11.5,6.6){$\phi(O_2)$}
+\put(28.5,6.6){$\phi(O_n)$}
+\put(7,16.5){\footnotesize{$\psi(X_1{,}X_2)$}}
+\put(15,16.5){\footnotesize{$\psi(X_2{,}X_3)$}}
+\put(22.5,16.5){\footnotesize{$\psi(X_{n{-}1}{,}X_n)$}}
+\put(5.2,10.6){\footnotesize{$\psi(X_1{,}O_1)$}}
+\put(13.2,10.6){\footnotesize{$\psi(X_2{,}O_2)$}}
+\put(30.2,10.6){\footnotesize{$\psi(X_n{,}O_n)$}}
+\end{picture}
+
+\end{figure}
+
+%\includegraphics[width=\textwidth]{dummy.pdf}
 

### 2012-10-12 docs/tex/draft.tex
Minor formatting changes
Modifications for abstract

@@
 \title{Pareto-Optimal Multistate Protein Design using Tree Graphical Models}
-\author{Subhodeep Moitra, CMU}
+\author{Subhodeep Moitra \\ {\tt subhodee@andrew.cmu.edu}}
 
@@
 \begin{abstract}
-To be added shortly
+The problem of protein design is to find a protein sequence which folds into a given protein structure.
+Multistate protein design generalizes this to find a protein sequence which folds into multiple protein states.
+This involves finding a sequence that is not only stable in one state but in the other state as well.
+This problem can be posed as a multi-objective optimization problem.
+By making suitable modelling approximations using graphical models, I propose to solve this multi-state design problem that allows the discovery of a set of sequences that are pareto optimal.
+Pareto-optimality in this context refers to the property that it is not possible to improve stability in one state without hurting the stability of the other state.
+Theory  , experiments and conclusions will be presented.
 \end{abstract}
+
+\clearpage
+\newpage
 
@@
 
+\pagebreak
+
 \subsection{Convex combinations of Energy Functions}
-
-\begin{equation}
-    \label{simple_equation}
-    \alpha = \sqrt{ \beta }
-\end{equation}
-
+To be added shortly
 %\begin{figure}

### 2012-10-12 docs/tex/compile_tex.sh
Some changes to Chain structured graphical model

@@ -14,6 +14,9 @@ bibtex ${fbase}.aux
 pdflatex ${fn}
 pdflatex ${fn}
 
+# Display the generated pdf file
+test -e ${fbase}.pdf && ( evince ${fbase}.pdf & ) || ( echo \
+  "File ${fbase}.pdf does not exist" >&2 )
 
 
 

### 2012-10-12 docs/tex/draft.tex
Some changes to Chain structured graphical model
Added section on CMRF

@@
 The CMRF replaces the Conditional probability tables(CPTs) present in the HMM with node and edge potentials.
-The following figure describes the CMRF.
+Figure~\ref{fig:CMRF} describes a CMRF.
+The max-product algorithm in the case of the CMRF is actually very similar to the dynamic programming solution of the HMM.
 
@@
 \end{picture}
+\label{fig:CMRF}
+\caption{A chain structured Markov Random Field. Probability of a sequence assignment is defined as a function of the node and edge potentials}
+\end{figure}
 
-\end{figure}
+\pagebreak
+
+The probability of the a sequence assignment is defined as follows :
+\begin{align*}
+P(X_1,X_2,\dotsc,X_n) &= \frac{1}{Z}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} \\
+\text{where} \quad Z &= \sum_{X_1,X_2,\dotsc,X_n}\left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\end{align*}
 
@@
 
-\pagebreak
+$Z$ is a normalizing constant known as the partition function.
+The value of $Z$ can be calculated efficiently using the sum-product algorithm.
+To calculate the maximum probability assignment $X^*$, we don't even need $Z$.
+$X^*$ can be evaluated using a dynamic programming formulation known as the max product algorithm.
+\begin{align*}
+X^* &= \argmax_{\mathbf{X}}P(\mathbf{X})\\
+ &= \argmax_{X_1,X_2,\dotsc,X_n}\frac{1}{Z}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} \\
+ &= \argmax_{X_1,X_2,\dotsc,X_n}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}
+\end{align*}
+
+ The main steps in the max-product algorithm are :
+
+\vspace*{0.2in}
+
+\textit{Calculation of the Optimal Probability value:}
+\[
+  V(i,k) = \left\{
+  \begin{array}{l l}
+    \phi(\mathbf{O_1})\psi(\mathbf{O_1},X_1{=}k)\phi(X_1{=}k) & \quad \text{if $i=1$}\\
+    \phi(\mathbf{O_i})\phi(X_i{=}k)\psi(\mathbf{O_i},X_i{=}k)\max_{j}\{\psi(X_{i}{=}k,X_{i-1}{=}j)V(i-1,j)\} & \quad \text{if $2\leq i\leq n$}\\
+  \end{array} \right.
+\]
+
+\textit{Retrieval of the optimal sequence corresponding to the optimal probability value:}
+\begin{align*}
+X^\star_n &= \argmax_{k}V(n,k) \\
+X^\star_{i} &= Ptr(X^\star_{i+1},i+1) & \quad \text{if $1\leq i \leq n-1 $}
+\end{align*}
+
+Here, $V(i,k)$ corresponds to the maximum functional value attainable by examining the sequence from $X_{1:i}$ and the observations from $\mathbf{O_{1:i}}$.
+Also, $X^\star_{1:n}$ corresponds to the optimal sequence assignment.
+Pointers are stored to each of the previous variables during the calculation of $V(i,k)$.
+These pointers are used to retrieve the optimal sequence assignment.
 

### 2012-10-12 docs/tex/draft.tex
Added section on multistate protein design

@@
 
-\subsection{Convex combinations of Energy Functions}
-To be added shortly
-%\begin{figure}
-%    \centering
-%    \includegraphics[width=3.0in]{myfigure}
-%    \caption{Simulation Results}
-%    \label{simulationfigure}
-%\end{figure}
+\subsection{Multistate protein design}
+The problem of protein structure prediction is finding the lowest energy structure given a particular sequence.
+Protein design is known as the inverse protein structure prediction problem and involves finding the lowest energy sequence for a given structure.
+Multistate protein design generalizes protein design by asking to find a sequence that is stable for multiple structures.
+This is often posed as a  multi-objective optimization problem \ref{}.
+
+\paragraph*{} Suppose, $E_A(X)$ and $E_B(X)$ are the energies of a sequence in state A and state B respectively.
+Then one way of doing multistate design would be to find $X^* = \arg \min_{X} E_A(X) + E_B(X)$.
+This would be the sequence that is likely to be stable under both state A and state B.
+Building on this idea, we might prefer to find a sequence $X^*$ that is more stable in one form than the other.
+We can thus solve for $X^* = \arg \min_{X} \theta_AE_A(X) + \theta_BE_B(X)$ where $\theta_A+\theta_B=1$ and $\theta_A,\theta_B \geq 0$.
+This sequence might be different from the solution to the previous optimization problem.
+By trying different combinations of $\theta_A$ and $\theta_B$ it is possible to find a number of different low energy sequences.
+However, it is not possible to try all possible values of $\theta$, nor is it possible to know which ones to pick to get a variety of different sequences.
+
+\paragraph*{} We will present an algorithm which finding optimal sequences for all combinations of $\theta_A$ and $\theta_B$.
+This is known as the pareto optimal frontier of the sequence space.
+See Figure~\ref{fig:pareto1} for an illustration of the pareto-optimal frontier.
+This frontier is pareto-optimal in the sense that it is not possible to unilaterally improve the energy function value of one state without adversely affecting the energy value of the other state.
+Thus each sequence on the frontier is optimal for certain sets of $\theta_A$ and $\theta_B$.
+
+\paragraph*{} Convex hull procedure
 

### 2012-12-06 docs/tex/draft.tex
Added pseudo code skeleton Added Tree MRF diagram Added section on what happens when function is not normalized Fixed some typos Added diagrams on convex hull illustration.

@@
 \usepackage{color}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
 
@@
 \end{abstract}
-
-\clearpage
-\newpage
 
@@
 \begin{align*}
-P(X_1,X_2,\dotsc,X_n) &= \frac{1}{Z}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} \\
-\text{where} \quad Z &= \sum_{X_1,X_2,\dotsc,X_n}\left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+P(X_1,X_2,\dotsc,X_n) &= \frac{1}{Z}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_i})\psi(\mathbf{O_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} \\
+\text{where} \quad Z &= \sum_{X_1,X_2,\dotsc,X_n}\left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_i})\psi(\mathbf{O_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
 \end{align*}
@@
 X^* &= \argmax_{\mathbf{X}}P(\mathbf{X})\\
- &= \argmax_{X_1,X_2,\dotsc,X_n}\frac{1}{Z}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} \\
- &= \argmax_{X_1,X_2,\dotsc,X_n}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_1})\psi(\mathbf{O_1},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}
+ &= \argmax_{X_1,X_2,\dotsc,X_n}\frac{1}{Z}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_i})\psi(\mathbf{O_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} \\
+ &= \argmax_{X_1,X_2,\dotsc,X_n}\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O_i})\psi(\mathbf{O_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}
 \end{align*}
@@
 
-\paragraph*{} We will present an algorithm which finding optimal sequences for all combinations of $\theta_A$ and $\theta_B$.
+\paragraph*{} We will present an algorithm for finding optimal sequences for all combinations of $\theta_A$ and $\theta_B$.
 This is known as the pareto optimal frontier of the sequence space.
-See Figure~\ref{fig:pareto1} for an illustration of the pareto-optimal frontier.
+See Figure~\ref{fig:cvxfront} for an illustration of the pareto-optimal frontier.
 This frontier is pareto-optimal in the sense that it is not possible to unilaterally improve the energy function value of one state without adversely affecting the energy value of the other state.
@@
 
-\paragraph*{} Convex hull procedure
+\subsubsection*{Quick hull procedure}
+The Quick hull procedure is used to find all the points on the pareto optimal frontier.
+The quick hull is traditionally used to find the convex hull of a set of points.
+In our case we are only interested in the lower left convex frontier.
+The algorithm in pseudo code is described in Algorithm~\ref{algo:quickhull1}.
+
+\begin{algorithm}
+\caption{Quick Hull} \label{algo:quickhull1}
+\begin{algorithmic}
+\For{each position $i$}
+	\For{each rotamer $r$ at $i$}
+		\For{each candidate rotamer $t$ at $i$}
+			\State $X \gets E_i(r_i^r) - E_i(r_i^t)$
+			\ForAll{positions $j \not= i$} 
+				\ForAll{rotamers $u$ at position $j$}
+					\State $Y = \min_u\left[E_{ij}(r_i^r,r_j^u) - E_{ij}(r_i^t,r_j^u) \right]$
+				\EndFor
+				\State $X \gets X + Y$
+			\EndFor
+			\If{$X>0$}
+				\State eliminate $r_i^r$
+				\State break
+			\EndIf
+		\EndFor
+	\EndFor
+\EndFor
+\end{algorithmic}
+\end{algorithm}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{pics/convexfront.pdf}
+    \caption{This figure illustrates the concept of a pareto-optimal convex frontier.
+The red curve is defined as the pareto-frontier and represents the tradeoff between $E_A(X)$ and $E_B(X)$.
+P1 and P2 are two points on the pareto frontier.
+P1 is better in terms of Energy A and P2 is better in terms of Energy B but not vice versa, thus showing a trade-off.
+The region D2(blue rectangle) is strictly dominated by P2, the region D1(green rectangle) is dominated by P1 and the region D12(yellow) is dominated by both P1 and P2.
+Domination means that points in these regions are clearly worse in both Energy A and Energy B.
+Also, notice the perpendicular drawn from the green dashed line and blue dashed line meets the pareto-frontier at P1 and P2 respectively.}
+    \label{fig:cvxfront}
+\end{figure}
+
+\pagebreak
+
+\subsection{Convex combinations of CMRFs}
+
+\begin{figure}[h]
+% selecting unit length
+\centering
+% used for centering Figure
+\setlength{\unitlength}{0.3cm}
+\begin{picture}(40,25)(0,0)
+\multiput(5,15)(8,0){2}{\circle{4}}
+\multiput(5,7)(8,0){2}{\circle{4}}
+\multiput(5,7)(8,0){2}{\circle{3.7}}
+\multiput(5,23)(8,0){2}{\circle{4}}
+\multiput(5,23)(8,0){2}{\circle{3.7}}
+\thicklines
+\multiput(7,15)(8,0){2}{\line(1,0){4}}
+\multiput(20,15)(1,0){4}{\circle*{0.3}}
+\multiput(5,13)(8,0){2}{\line(0,-1){4}}
+\multiput(5,17)(8,0){2}{\line(0,1){4}}
+\put(24,15){\line(1,0){4}}
+\put(30,13){\line(0,-1){4}}
+\put(30,17){\line(0,1){4}}
+\thinlines
+\put(30,15){\circle{4}}
+\put(30,7){\circle{4}}
+\put(30,7){\circle{3.7}}
+\put(30,23){\circle{4}}
+\put(30,23){\circle{3.7}}
+\put(3.5,14.6){$\phi(X_1)$}
+\put(11.5,14.6){$\phi(X_2)$}
+\put(28.5,14.6){$\phi(X_n)$}
+\put(3.5,6.6){$\phi(O^A_1)$}
+\put(11.5,6.6){$\phi(O^A_2)$}
+\put(28.5,6.6){$\phi(O^A_n)$}
+\put(3.5,22.6){$\phi(O^B_1)$}
+\put(11.5,22.6){$\phi(O^B_2)$}
+\put(28.5,22.6){$\phi(O^B_n)$}
+\put(7,16.5){\footnotesize{$\psi(X_1{,}X_2)$}}
+\put(15,16.5){\footnotesize{$\psi(X_2{,}X_3)$}}
+\put(22.5,16.5){\footnotesize{$\psi(X_{n{-}1}{,}X_n)$}}
+\put(5.2,10.6){\footnotesize{$\psi(X_1{,}O^A_1)$}}
+\put(13.2,10.6){\footnotesize{$\psi(X_2{,}O^A_2)$}}
+\put(30.2,10.6){\footnotesize{$\psi(X_n{,}O^A_n)$}}
+\put(5.2,18.6){\footnotesize{$\psi(X_1{,}O^B_1)$}}
+\put(13.2,18.6){\footnotesize{$\psi(X_2{,}O^B_2)$}}
+\put(30.2,18.6){\footnotesize{$\psi(X_n{,}O^B_n)$}}
+\end{picture}
+\label{fig:CMRF2}
+\caption{A Tree structured Markov Random Field that is used for modelling multi-state protein design.}
+\end{figure}
+
+\pagebreak
+
+\subsection{Convex Hull Procedure on CMRFs}
+
+\subsubsection{Normalizing CMRFs}
+So far we have been assuming that the $E_A(X)$ and $E_B(X)$ lies in the positive quadrant.
+However this assumption need not not be true under our CMRF model.
+This is because it is possible that $f(X) > 1$ and conversely $-\log f(X) < 0$.
+This can lead to the energy function values being negative as shown in Figure~\ref{fig:cvxfront2}.
+We can potentially fix this be normalizing energy the Likelihood function with the partition function Z.
+The partitions can be efficiently calculated using the sum-product algorithm for trees.
+However we will show that this step is not needed.
+
+For e.g. let the partition functions for each of the states be.
+\[
+Z_A = \sum_{X_1,X_2,\dotsc,X_n}\left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\]
+and 
+\[
+Z_B = \sum_{X_1,X_2,\dotsc,X_n}\left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^B_i})\psi(\mathbf{O^B_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\]
+\\
+If instead of the regular energy function 
+\[
+E_A(X) = -\log  \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\]
+we normalized the energy function with $Z_A$
+\[
+\begin{split}
+E_A'(X) &= -\log  \frac{1}{Z_A} \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]\\
+ &= -\log   \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right] + \log Z_A \\
+	&= E_A(X) + \log Z_A
+\end{split}
+\]
+\\
+Then optimizing the multistate Energy function would amount to : 
+\[
+\begin{split}
+X^{\dagger} &= \arg \min_X \left( \theta_A E_A'(X)  + \theta_B E_B'(X) \right) \\
+ &= \arg \min_X \left( \theta_A E_A(X) + \theta_B E_B(X) + \theta_A \log Z_A + \theta_B \log Z_B \right) \\
+ &= \arg \min_X \left( \theta_A E_A(X) + \theta_B E_B(X) \right) \\
+  &= X^*
+\end{split}
+\]
+
+This shows that normalizing the energy function values by $Z_A$ and $Z_B$ is not necessary in order to find point on the pareto optimal frontier.
+This is further demonstrated in Figure~\ref{fig:cvxfront2}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{pics/convexfront2.pdf}
+    \caption{This figure illustrates the issue with normalizing.
+When dealing with negative log likelihood values, it is possible that pareto frontier may not lie in the positive quadrant.
+This however does not imply that the method fails.
+Normalizing with the partition function merely corresponds to translating the points in the energy space.
+This transformation keeps the optimal solution invariant.
+}
+    \label{fig:cvxfront2}
+\end{figure}
+\pagebreak
 

### 2012-12-06 docs/tex/draft.tex
Completed psuedo-code of quickhull
Added clarification on energy convex combination Added references. Described quickhull. Started writing CMRF convex.

@@
 Multistate protein design generalizes protein design by asking to find a sequence that is stable for multiple structures.
-This is often posed as a  multi-objective optimization problem \ref{}.
+This is often posed as a  multi-objective optimization problem~\cite{Zheng2009}.
 
@@
 
-\subsubsection*{Quick hull procedure}
-The Quick hull procedure is used to find all the points on the pareto optimal frontier.
-The quick hull is traditionally used to find the convex hull of a set of points.
-In our case we are only interested in the lower left convex frontier.
-The algorithm in pseudo code is described in Algorithm~\ref{algo:quickhull1}.
-
-\begin{algorithm}
-\caption{Quick Hull} \label{algo:quickhull1}
-\begin{algorithmic}
-\For{each position $i$}
-	\For{each rotamer $r$ at $i$}
-		\For{each candidate rotamer $t$ at $i$}
-			\State $X \gets E_i(r_i^r) - E_i(r_i^t)$
-			\ForAll{positions $j \not= i$} 
-				\ForAll{rotamers $u$ at position $j$}
-					\State $Y = \min_u\left[E_{ij}(r_i^r,r_j^u) - E_{ij}(r_i^t,r_j^u) \right]$
-				\EndFor
-				\State $X \gets X + Y$
-			\EndFor
-			\If{$X>0$}
-				\State eliminate $r_i^r$
-				\State break
-			\EndIf
-		\EndFor
-	\EndFor
-\EndFor
-\end{algorithmic}
-\end{algorithm}
-
 \begin{figure}[h!]
@@
 
+\subsubsection*{Quick hull procedure}
+The Quick hull procedure is used to find all the points on the pareto optimal frontier.
+The quick hull is traditionally used to find the convex hull of a set of points.
+In our case we are only interested in the lower left convex frontier similar to ~\cite{Zheng2009}.
+The algorithm in pseudo code is described in Algorithm~\ref{algo:quickhull1}.
+The basic idea behind the algorithm is to find a pair of points on the pareto optimal frontier.
+Find the point farthest from the line joining the two points and then to repeat the procedure with this newly found point and the previous two points.
+When there are no newer points to find, all points on the pareto optimal frontier have been discovered.
+
+\begin{algorithm}
+\caption{Quick Hull} \label{algo:quickhull1}
+\begin{algorithmic}
+\State Initialize queue $Q \gets \emptyset$
+\State Find $X_A \gets \arg \min_X E_A(X) $
+\State Find $X_B \gets \arg \min_X E_B(X) $
+\State Enqueue $(X_A,X_B)$
+\State $C_H \gets \{X_A,X_B\}$
+\Repeat
+	\State Dequeue from Q a pair $(X_1,X_2)$ and assert($E_A(X_1) < E_A(X_2)$)
+	\State $m \gets \frac{E_B(X_1)-E_B(X_2)}{E_A(X_1)-E_A(X_2)}$
+	\State $\theta_A \gets \frac{-m}{1-m}$
+	\State $\theta_B \gets \frac{1}{1-m}$
+	\State Find $X_{AB} \gets \arg \min_X \theta_A E_A(X) + \theta_B E_B(X) $
+	\If{$X_{AB} \neq X_A$ and $X_{AB} \neq X_B$ }
+		\State $C_H \gets C_H \cup \{X_{AB}\}$
+		\State Enqueue $(X_A,X_{AB})$ and $(X_{AB},X_B)$
+	\EndIf
+\Until{Q is empty}
+\State \Return $C_H$
+\end{algorithmic}
+\end{algorithm}
+
 \subsection{Convex combinations of CMRFs}
+For multi-state protein design we can combine the energy functions as follows.
+\[
+E_{AB}(X) = \theta_A E_A(X) + \theta_B E_B(X) \quad \text{where} \quad \theta_A,\theta_B \geq 0 \text{ and } \theta_A + \theta_B =1
+\]
+Here $E_A(X)$ is any energy function mapping sequence to energy values for state A. The lower the energy the better the fit for sequence X to state A. $E_A(X)$ can also be written as a CMRF.
+\[
+E_A(X) = -\log  \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\]
+Note that $-\log$ is to formulate the problem as a minimization problem so that it agrees with the energy concept. However the optimal sequence remain the same. 
+\[
+\begin{split}
+X_A &= \arg \min_X -\log  \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]\\
+&= \arg \max_X \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\end{split}
+\]
+Taking the $-\log$ also offers benefits as it lets us combine convex combinations of CMRF energy functions and gives rise to new graphical models called Tree MRFS (TMRFs).
+These TMRFs are conceptually very similar to CMRFs and possess the properties of efficient inference and decoding.
 
-\begin{figure}[h]
+\begin{align*}
+X_A &= \arg \min_X -\log  \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+\end{align*}
+
+\begin{figure}[h!]
 % selecting unit length
@@
 
+\subsubsection{Clarification on Energy Convex Combination}
+From the discussion in the beginning, the reader may have been lead to believe that we are taking convex combinations of the likelihood of multiple states and then maximizing this likelihood.
+While in spirit this is true, however practically we are optimizing something else.
+Let $L(X|O^A)$  and $L(X|O^B)$ be the likelihoods of the sequences under state A and state B respectively.
+The reader may have been lead to believe that we are solving the following problem
+\begin{equation}
+\label{eqn:clarif1}
+X^* = \arg \max_X \left( \theta_A L(X|O^A) + \theta_B L(X|O^B) \right)
+\end{equation}
+But in, reality we are solving : 
+\begin{align}
+X^* &= \arg \min_X \left( -\theta_A \log L(X|O^A) - \theta_B \log L(X|O^B) \right) \\
+&= \arg \min_X \left( \log \frac{1}{L(X|O^A)^{\theta_A} L(X|O^B)^{\theta_B}} \right) \\
+&= \arg \max_X \left( L(X|O^A)^{\theta_A} L(X|O^B)^{\theta_B} \right)
+\label{eqn:clarif2} 
+\end{align}
+
+Clearly Eqn~\ref{eqn:clarif2} is not the same as Eqn~\ref{eqn:clarif1}.
+One is minimizing the arithmetic mean, while the other is minimizing the geometric mean.
+It remains to be seen what the semantic consequence of this mathematical detail is.
+
 \section{Conclusion}

### 2012-12-06 docs/tex/draft.tex
Added the meat of how to max decoding on Tree MRFs.

@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
 
@@
 \begin{align*}
-X_A &= \arg \min_X -\log  \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]
+X_A &= \arg \min_X E_A(X) \\
+&= \arg \min_X -\log  \left[{\prod_{i=1}^{n}{\phi(X_i)\phi(\mathbf{O^A_i})\psi(\mathbf{O^A_i},X_i})\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}}\right]\\
+&= \arg \min_X -\left[{\sum_{i=1}^{n} \log\phi(X_i)+\sum_{i=1}^{n} \log\phi(\mathbf{O^A_i})+\sum_{i=1}^{n} \log\psi(\mathbf{O^A_i},X_i)+ \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}}\right] \\
+&= \arg \min_X -\left[{\sum_{i=1}^{n} \log\phi(X_i)+\sum_{i=1}^{n} \log\psi(\mathbf{O^A_i},X_i)+ \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}}\right]
 \end{align*}
+
+Similarly,
+\begin{align*}
+X_B &= \arg \min_X E_B(X) \\
+&= \arg \min_X -\left[{\sum_{i=1}^{n} \log\phi(X_i)+\sum_{i=1}^{n} \log\psi(\mathbf{O^B_i},X_i)+ \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}}\right]
+\end{align*}
+\\
+Taking the convex combination of the energy functions we get 
+\begin{align*}
+X_{AB} &= \arg \min_X \theta_A E_A(X) + \theta_B E_B(X) \\
+\text{s.t} &\quad  \theta_A + \theta_B  = 1 \\
+& \theta_A, \theta_B \geq 0
+\end{align*}
+Expanding $E_A(X)$ and $E_B(X)$ we get, 
+\begin{align*}
+X_{AB} &= \arg \min_X -\theta_A \left[{\sum_{i=1}^{n} \log\phi(X_i)+\sum_{i=1}^{n} \log\psi(\mathbf{O^A_i},X_i)+ \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}}\right] \\
+&-\theta_B \left[{\sum_{i=1}^{n} \log\phi(X_i)+\sum_{i=1}^{n} \log\psi(\mathbf{O^B_i},X_i)+ \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}}\right] \\
+&= \arg \min_X -\left( \sum_{i=1}^{n}{\log\phi(X_i)} + \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}\right)\\ 
+ &-\left( \sum_{i=1}^{n}{(\theta_A\log\psi(\mathbf{O^A_i},X_i) + \theta_B\log\psi(\mathbf{O^B_i},X_i))} \right) \\
+ &= \arg \min_X -\left( \sum_{i=1}^{n}{\log\phi(X_i)} + \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}\right)\\ 
+ &-\left( \sum_{i=1}^{n}{(\log\psi(\mathbf{O^A_i},X_i)^{\theta_A} + \log\psi(\mathbf{O^B_i},X_i)^{\theta_B})} \right)
+\end{align*}
+
+This looks very much like decoding on a CMRF with the added change that the emission edge potential values are exponentiated.
+In fact this distribution factorizes according to the TMRF shown in Figure~\ref{fig:CMRF2}.
 
@@
 \put(22.5,16.5){\footnotesize{$\psi(X_{n{-}1}{,}X_n)$}}
-\put(5.2,10.6){\footnotesize{$\psi(X_1{,}O^A_1)$}}
-\put(13.2,10.6){\footnotesize{$\psi(X_2{,}O^A_2)$}}
-\put(30.2,10.6){\footnotesize{$\psi(X_n{,}O^A_n)$}}
-\put(5.2,18.6){\footnotesize{$\psi(X_1{,}O^B_1)$}}
-\put(13.2,18.6){\footnotesize{$\psi(X_2{,}O^B_2)$}}
-\put(30.2,18.6){\footnotesize{$\psi(X_n{,}O^B_n)$}}
+\put(5.2,10.6){\footnotesize{$\psi^{\theta_A}(X_1{,}O^A_1)$}}
+\put(13.2,10.6){\footnotesize{$\psi^{\theta_A}(X_2{,}O^A_2)$}}
+\put(30.2,10.6){\footnotesize{$\psi^{\theta_A}(X_n{,}O^A_n)$}}
+\put(5.2,18.6){\footnotesize{$\psi^{\theta_B}(X_1{,}O^B_1)$}}
+\put(13.2,18.6){\footnotesize{$\psi^{\theta_B}(X_2{,}O^B_2)$}}
+\put(30.2,18.6){\footnotesize{$\psi^{\theta_B}(X_n{,}O^B_n)$}}
 \end{picture}
 \label{fig:CMRF2}
-\caption{A Tree structured Markov Random Field that is used for modelling multi-state protein design.}
+\caption{A Tree structured Markov Random Field that is used for modelling multi-state protein design.
+This Tree structured MRF results from taking a convex combination $\theta_A E_A(X) + \theta_B E_B(X)$ of the log likelihoods of CMRFs for State A and State B }
 \end{figure}
@@
 
-\subsection{Convex Hull Procedure on CMRFs}
+\subsection{Max-Decoding on TMRFs}
+In the previous section we showed how the convex combination of negative log likelihoods of CMRFs can be used to construct a TMRF.
+Here we will show how efficient decoding can be performed on the TMRF using the similar variant of Viterbi decoding.
+
+\begin{align*}
+X_{AB} &= \argmin_X \left( \theta_A E_A(X) + \theta_B E_B(X)\right) \\
+&= \arg \min_X \left( -\sum_{i=1}^{n}{\log\phi(X_i)} + \sum_{i=1}^{n-1}{\log\psi(X_i,X_{i+1})}\right)\\ 
+ &-\left( \sum_{i=1}^{n}{(\log\psi(\mathbf{O^A_i},X_i)^{\theta_A} + \log\psi(\mathbf{O^B_i},X_i)^{\theta_B})} \right) \\
+ &= \argmax_{X}\log\left( \prod_{i=1}^{n}{\phi(X_i)\psi(\mathbf{O^A_i},X_i)^{\theta_A} \psi(\mathbf{O^B_i},X_i)^{\theta_B}}\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}\right) \\
+ &= \argmax_{X}\left( \prod_{i=1}^{n}{\phi(X_i)\psi(\mathbf{O^A_i},X_i)^{\theta_A} \psi(\mathbf{O^B_i},X_i)^{\theta_B}}\prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})}\right)
+\end{align*}
+
+ The main steps in the max-product algorithm are :
+
+\vspace*{0.2in}
+
+\textit{Calculation of the Optimal Probability value:}
+\[
+  V(i,k) = \left\{
+  \begin{array}{l l}
+    \psi(\mathbf{O^A_1},X_1{=}k)^{\theta_A}\psi(\mathbf{O^B_1},X_1{=}k)^{\theta_B}\phi(X_1{=}k) & \quad \text{if $i=1$}\\
+    \phi(X_i{=}k)\psi(\mathbf{O^A_i},X_i{=}k)^{\theta_A}\psi(\mathbf{O^B_1},X_1{=}k)^{\theta_B}\cdots\\
+    \max_{j}\{\psi(X_{i}{=}k,X_{i-1}{=}j)V(i-1,j)\} & \quad \text{if $2\leq i\leq n$}\\
+  \end{array} \right.
+\]
+
+\textit{Retrieval of the optimal sequence corresponding to the optimal probability value:}
+\begin{align*}
+X^\star_n &= \argmax_{k}V(n,k) \\
+X^\star_{i} &= Ptr(X^\star_{i+1},i+1) & \quad \text{if $1\leq i \leq n-1 $}
+\end{align*}
+
+Here, $V(i,k)$ corresponds to the maximum functional value attainable by examining the sequence from $X_{1:i}$ and the observations from $\mathbf{O^A_{1:i}}$ and $\mathbf{O^B_{1:i}}$.
+Also, $X^\star_{1:n}$ corresponds to the optimal sequence assignment.
+Pointers are stored to each of the previous variables during the calculation of $V(i,k)$.
+These pointers are used to retrieve the optimal sequence assignment.
+
+\subsection{Convex Hull Procedure on TMRFs}
+In the previous section we showed how to find $X_{AB} = \argmin_X \left( \theta_A E_A(X) + \theta_B E_B(X)\right)$, for any convex combination with $\theta_A$ and $\theta_B$.In this section, we will put it all together to define a comprehensive algorithm that can be used for finding the entire pareto frontier for sequence design tasks.
 

### 2012-12-06 docs/tex/draft.tex
Created comprehensive algorithm for tying all parts together

@@
 \subsection{Convex Hull Procedure on TMRFs}
-In the previous section we showed how to find $X_{AB} = \argmin_X \left( \theta_A E_A(X) + \theta_B E_B(X)\right)$, for any convex combination with $\theta_A$ and $\theta_B$.In this section, we will put it all together to define a comprehensive algorithm that can be used for finding the entire pareto frontier for sequence design tasks.
+In the previous section we showed how to find $X_{AB} = \argmin_X \left( \theta_A E_A(X) + \theta_B E_B(X)\right)$, for any convex combination with $\theta_A$ and $\theta_B$.In this section, we will put it all together in Algorithm~\ref{algo:pareto1} and Algorithm~\ref{algo:pareto2} to define a comprehensive algorithm that can be used for finding the entire pareto frontier for sequence design tasks.
+
+\begin{algorithm}
+\caption{Pareto-Optimal MultiState Protein Design} \label{algo:pareto1}
+\begin{algorithmic}[1]
+\Procedure{Pareto-Frontier}{$\phi,\psi,O^A,O^B$}\Comment{Find Pareto frontier}
+	\State Initialize queue $Q \gets \emptyset$
+	\State $X_A \gets$ \Call{CMRF-Decode}{$\phi,\psi,O^A$}  \Comment{$\argmin_X E_A(X)$}
+	\State $X_B \gets$ \Call{CMRF-Decode}{$\phi,\psi,O^B$}  \Comment{$\argmin_X E_B(X)$}
+	\State Enqueue $(X_A,X_B)$
+	\State $C_H \gets \{X_A,X_B\}$
+	\Repeat
+		\State Dequeue from Q a pair $(X_1,X_2)$ and assert($E_A(X_1) < E_A(X_2)$)
+		\State $m \gets \frac{E_B(X_1)-E_B(X_2)}{E_A(X_1)-E_A(X_2)}$
+		\State $\theta_A \gets \frac{-m}{1-m}$
+		\State $\theta_B \gets \frac{1}{1-m}$
+		\State $X_{AB} \gets $\Call{TMRF-Decode}{$\phi,\psi,O^A,O^B,\theta_A,\theta_B$}
+		\Statex \Comment{$\argmin_X \theta_A E_A(X) + \theta_B E_B(X)$}		
+		\If{$X_{AB} \neq X_A$ and $X_{AB} \neq X_B$ }
+			\State $C_H \gets C_H \cup \{X_{AB}\}$
+			\State Enqueue $(X_A,X_{AB})$ and $(X_{AB},X_B)$
+		\EndIf
+	\Until{Q is empty}
+	\State \Return $C_H$
+\EndProcedure
+\Statex 
+\Procedure{CMRF-Decode}{$\phi,\psi,O^A$}\Comment{Max decode CMRF}
+	\State Allocate $V[N][K]$ \Comment{N is length of seq,K is numAA }
+	\State Allocate $X_A[N]$	
+	\For{$k \gets 1,K$}
+		\State $V[1][k] \gets \psi(\mathbf{O^A_1},X_1{=}k)\phi(X_1{=}k)$
+	\EndFor	
+	\For{$i \gets 1,N$}
+		\For{$k \gets 1,K$}
+			\State $V[i][k] \gets \phi(X_i{=}k)\psi(\mathbf{O^A_i},X_i{=}k) \max_{j}\{\psi(X_{i}{=}k,X_{i-1}{=}j)V(i-1,j)\}$
+			\State $Ptr[i][k] \gets \argmax_{j}\{\psi(X_{i}{=}k,X_{i-1}{=}j)V(i-1,j)\} $
+		\EndFor		
+	\EndFor
+	\Statex \Comment{Now retrieve the pointers}
+	\State $X_A[N] \gets \argmax_{k}V[N][k]$ 
+	\For{$i \gets N-1,1$}
+		\State	$X_A[i] \gets Ptr[i+1][X_A[i+1]]$
+	\EndFor
+	\State \bf{return} $X_A$
+\EndProcedure
+\algstore{paretobreak}
+\end{algorithmic}
+\end{algorithm}
+
+\begin{algorithm}
+\caption{Pareto continued} \label{algo:pareto2}
+\begin{algorithmic}[1]
+\algrestore{paretobreak}
+
+\Procedure{TMRF-Decode}{$\phi,\psi,O^A,O^B,\theta_A,\theta_B$}\Comment{Max Decode TMRF}
+	\State Allocate $V[N][K]$ \Comment{N is length of seq,K is numAA }
+	\State Allocate $X_{AB}[N]$	
+	\For{$k \gets 1,K$}
+		\State $V[1][k] \gets \psi(\mathbf{O^A_1},X_1{=}k)^{\theta_A}\psi(\mathbf{O^B_1},X_1{=}k)^{\theta_B}\phi(X_1{=}k)$
+	\EndFor	
+	\For{$i \gets 1,N$}
+		\For{$k \gets 1,K$}
+			\State $V[i][k] \gets \phi(X_i{=}k)\psi(\mathbf{O^A_i},X_i{=}k)^{\theta_A}\psi(\mathbf{O^B_1},X_1{=}k)^{\theta_B}
+    \max_{j}\{\psi(X_{i}{=}k,X_{i-1}{=}j)V(i-1,j)\}$
+			\State $Ptr[i][k] \gets \argmax_{j}\{\psi(X_{i}{=}k,X_{i-1}{=}j)V(i-1,j)\} $
+		\EndFor		
+	\EndFor
+	\Statex \Comment{Now retrieve the pointers}
+	\State $X_{AB}[N] \gets \argmax_{k}V[N][k]$ 
+	\For{$i \gets N-1,1$}
+		\State	$X_{AB}[i] \gets Ptr[i+1][X_{AB}[i+1]]$
+	\EndFor
+	\State \bf{return} $X_{AB}$
+\EndProcedure
+
+\end{algorithmic}
+\end{algorithm}
+
+\pagebreak
 

### 2012-12-06 docs/tex/draft.tex
Introduction and Conclusion

@@
 This problem can be posed as a multi-objective optimization problem.
-By making suitable modelling approximations using graphical models, I propose to solve this multi-state design problem that allows the discovery of a set of sequences that are pareto optimal.
-Pareto-optimality in this context refers to the property that it is not possible to improve stability in one state without hurting the stability of the other state.
-Theory  , experiments and conclusions will be presented.
+By making suitable modelling approximations using graphical models we attempt to solve this optimization problem exactly and efficiently.
+We also show how to efficiently find all pareto-optimal sequences by enumerating the pareto-optimal frontier in sequence space.
+Thus, we make a novel contribution to the repertoire of algorithms for protein design.
 \end{abstract}
@@
 \section{Introduction}
-To be added shortly \cite{Ambroggio2006}.
+A major goal in medicine is to be able to design biological macromolecules that perform a certain desired biological function.
+These biological macromolecules called proteins are ubiquitously found in all living organisms.
+They perform a number of functions ranging from motion, structure, signalling, catalysis, etc.
+They are the molecular machines of the biological world.
+One of the primary reasons that proteins are able to perform so many diverse functions is that individual proteins can adopt multiple roles.
+They do so by being able to show great structural diversity.
+While it is certainly true that sequence determines structure, it is not true that it is a single static structure.
+Proteins can adopt multiple structures and conformations.
+Some proteins are particularly better than others in their ability to adopt distinct conformations.
+We refer to these proteins as multistate proteins.
+\\
+\\
+The problem of protein design is to find a protein sequence which folds into a given protein structure.
+Multistate protein design generalizes this to find a protein sequence which folds into multiple protein states.
+This involves finding a sequence that is not only stable in one state but in the other state as well.
+Biologist have previously designed proteins that can switch between folded and disordered states, between distinct folded states and between different aggregation states~\cite{Ambroggio2006}.
+These proteins switch under a variety of trigger mechanisms such as pH shifts, post-translational modification and ligand binding.
+However most efforts lead by biologists have heavily relied on detailed prior understanding of the system and trial and error.
+\\
+\\
+Computational attempts at multistate design have been stymied by the enormous complexity of the protein design space.
+Pierce et.al~\cite{Pierce2002} proved that protein design is NP-hard.
+Multistate design a generalization of protein design is by extension NP-hard as well.
+Previous attempts at multistate protein design have involved a lot of different strategies including Markov Chain Monte Carlo (MCMC)~\cite{Leaver-Fay2011,Ambroggio2006}, Genetic Algorithms, Dead End Elimination~\cite{Yanover2007} and even graphical models~\cite{Fromer2010}.The multistate applications have ranged from negative design, protein-protein binding, homodimerization, allosteric design, etc.
+Each of these methods make significant contributions to the field of multistate protein design.
+However each of them suffers from some drawback or the other.
+MCMC and GA methods are approximate and do not have any guarantees and are expensive.
+DEE is an exact method, but is very expensive to calculate and cannot be scaled up.
+BP using MRFs is the most elegant yet, but it also suffers from computational cost and scale-up issues.
+\\
+\\
+When designing algorithms, there is often a trade-off between accuracy and speed.
+In this paper, I will present a method that is fast and exact.
+It is very flexible allowing the user to supply various features.
+It also allows for full enumeration of the pareto-frontier of the sequence space.
+Unfortunately, these benefits come at the cost of accuracy and expressive power.
+Specifically, it makes an sequential dependency assumption between the amino acids of a protein.
+This is obviously not true since proteins are 3D structures and can fold back in order to interact with non-sequential parts of itself.
+While it is true that this is a limitation, it remains to be seen how much of a limitation it really is.
+Many chain methods such as HMMs are very successfully used within the biological community.
+\\
+\\
+Our method uses tree-structured MRFs(TMRFs)for efficient inference and decoding to find the minimum energy sequence for a particular state.
+We make a clever observation that minimum energy sequences of convex combinations of states can be found efficiently as well when using a TMRF.
+This unique property lets us fully enumerate the pareto-frontier of the sequence space thus identifying all the sequences that are strictly dominant to other sequences.
+Thus it lets us very quickly describe the sequence space.
+In the rest of the paper we will present the theory behind all the methods.
+We first introduce Hidden Markov Models (HMMs) for motivation.
+We then move onto chain structured CMRFs and talk about efficient max-decoding in these frameworks.
+We further generalize CMRFs to tree structured MRFs (TMRFs) and show how to do max decoding under this setting as well.
+Then we tie all the models together using the quick hull algorithm for enumerating the sequences in the pareto frontier.
+Finally, we close by talking about some mathematical subtleties of the proposed method.
 
@@
 \section{Conclusion}
-Write your conclusion here.
+We presented a method that is able to perform fast decoding and inference for finding minimum energy sequences for the purpose of protein design.
+We also show to find the pareto-frontier for the sequence space.
+The key idea behind this method is the use of Tree MRFs which allow fast Dynamic Programming for max-decoding.
 

### 2012-12-19 docs/tex/draft.tex
Experiments and simulations

@@
 \usepackage{algorithm}
+\usepackage{multirow}
+\usepackage{color}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
 
@@
 
+\subsection{Parameter Learning/Training}
+So far, we have assumed that the $\phi,\psi$ parameters were available to us.
+In this section we will describe how these weights are learned.
+How we can use these weights to for defining the joint TMRF.
+
+\subsubsection{Training an HMM}
+
+\subsubsection{Converting a HMM to CMRF}
+It is possible to reparameterize an HMM so that it now becomes a CMRF. The partition function $Z$ of the newly constructed CMRF equals 1 corresponding to a probability distribution., However there is
+
+\subsection{Combining the CMRF to a HMM}
+
+\section{Experiments}
+In this section we evaluate the performance of our algorithm on simulated as well as experimental data.
+
+\subsection{Simulation Experiments}
+We demonstrate how our algorithm performs on simulated data.
+The data was prepared by considering a protein of length 15.
+The sequence space was limited to an alphabet consisting of two arbitrary characters $\alpha,\beta$.
+The observation belong to a discrete space Helix, Beta and Loop which represented by H,B and L.
+The weights of the CMRF have weights $\phi,\psi$ as defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2}.
+We define two states State A and State B that corresponding to competing objectives for multistate design.
+State A is a Helix-Loop-Helix strand whereas State B is Sheet-Loop-Sheet strand.
+Figure~\ref{fig:simstates} shows a cartoon diagrams for our design problem.
+State A and State B have \texttt{HHHHLLLHHHH} and \texttt{BBBBLLLBBBB} secondary structure features respectively.
+These are the observations for the CMRF.
+\\
+\\
+Using the weight matrices we defined in Tables~\ref{tab:ps1} and~\ref{tab:psi2} we run CMRF-Decode to get $X_A$ which satisfies $X_A = \argmin_X E_A(X)$.
+We obtain $X_B$ similarly.
+Not surprisingly $X_A$ is $\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha$
+and $X_B$ is $\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta$.
+This is because the weights in the table are biased towards $\alpha$ emitting \texttt{H} and $\beta$ emitting \texttt{B}.
+Thus according to our model $X_A$ is the sequence that is most stable under state A, whereas $X_B$ is the sequence is most stable under state B.
+\\
+\\
+We then run the \texttt{Pareto-Frontier} algorithm using the same weights and observations for state A and state B to enumerate the entire pareto frontier.
+We number the points to show the path taken by the pareto frontier algorithm.
+We also calculate the energies of all the $2^15$ sequences possible under this restricted sequence model and plot the points.
+The \texttt{Pareto-Frontier} algorithm was able to find the pareto frontier in \textcolor{red}{N steps} steps.
+See Figure~\ref{fig:pareto-sim} for the trace taken by the algorithm.
+We also compare the running times as opposed to brute force and MCMC and compare the running times.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{cc|c|c|c|}
+\cline{3-5}
+& & \multicolumn{3}{c|}{Observations($\mathbf{O_i}$)} \\ \cline{3-5}
+& & Helix(H) & Beta(B) & Loop(L) \\ \cline{1-5}
+\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
+\multicolumn{1}{|c|}{$\alpha$} & 0.6 & 0.2 & 0.2      \\ \cline{2-5}
+\multicolumn{1}{|c|}{}                        &
+\multicolumn{1}{|c|}{$\beta$} & 0.2 & 0.6 & 0.2      \\ \cline{1-5}
+\end{tabular}
+\end{center}
+\caption{$\psi(X,\mathbf{O})$ contains the weights for edge potentials of sequence and features together }
+\label{tab:psi1}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{cc|c|c|}
+\cline{3-4}
+& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
+& & $\alpha$ & $\beta$  \\ \cline{1-4}
+\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
+\multicolumn{1}{|c|}{$\alpha$} & 0.7 & 0.3      \\ \cline{2-4}
+\multicolumn{1}{|c|}{}                        &
+\multicolumn{1}{|c|}{$\beta$} & 0.3 & 0.7      \\ \cline{1-4}
+\end{tabular}
+\end{center}
+\caption{$\psi(X_i,X_{i+1})$ contains the weights for the edge potentials of adjacent amino acids}
+\label{tab:psi2}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{cc|c|c|}
+\cline{3-4}
+& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
+& & $\alpha$ & $\beta$  \\ \cline{1-4}
+\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
+\multicolumn{1}{|c|}{$\alpha$} & 0.7 & 0.3      \\ \cline{2-4}
+\multicolumn{1}{|c|}{}                        &
+\multicolumn{1}{|c|}{$\beta$} & 0.3 & 0.7      \\ \cline{1-4}
+\end{tabular}
+\end{center}
+\caption{Time and steps taken on simulated data}
+\end{table}
+
+\section{Discussion}
+
 \subsubsection{Normalizing CMRFs}

### 2012-12-20 docs/tex/draft.tex
More on simulated data

@@
 We demonstrate how our algorithm performs on simulated data.
-The data was prepared by considering a protein of length 15.
+The data was prepared by considering a protein of length 12.
 The sequence space was limited to an alphabet consisting of two arbitrary characters $\alpha,\beta$.
@@
 Figure~\ref{fig:simstates} shows a cartoon diagrams for our design problem.
-State A and State B have \texttt{HHHHLLLHHHH} and \texttt{BBBBLLLBBBB} secondary structure features respectively.
+State A and State B have \texttt{HHHHLLLLHHHH} and \texttt{BBBBLLLLBBBB} secondary structure features respectively.
 These are the observations for the CMRF.
@@
 \\
-Using the weight matrices we defined in Tables~\ref{tab:ps1} and~\ref{tab:psi2} we run CMRF-Decode to get $X_A$ which satisfies $X_A = \argmin_X E_A(X)$.
+Using the weight matrices we defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2} we run CMRF-Decode to get $X_A$ which satisfies $X_A = \argmin_X E_A(X)$.
 We obtain $X_B$ similarly.
-Not surprisingly $X_A$ is $\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha$
-and $X_B$ is $\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta$.
+Not surprisingly $X_A$ is $\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha$
+and $X_B$ is $\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta$.
 This is because the weights in the table are biased towards $\alpha$ emitting \texttt{H} and $\beta$ emitting \texttt{B}.
@@
 We then run the \texttt{Pareto-Frontier} algorithm using the same weights and observations for state A and state B to enumerate the entire pareto frontier.
+This pareto frontier contains all the best sequences that any convex combination of $E_A(X)$ and $E_B(X)$ would give rise to.
 We number the points to show the path taken by the pareto frontier algorithm.
-We also calculate the energies of all the $2^15$ sequences possible under this restricted sequence model and plot the points.
+We also calculate the energies of all the $2^{12}$ sequences possible under this restricted sequence model and plot the points.
 The \texttt{Pareto-Frontier} algorithm was able to find the pareto frontier in \textcolor{red}{N steps} steps.
@@
 \end{table}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{pics/convexfront2.pdf}
+    \caption{ }
+    \label{fig:pareto-sim}
+\end{figure}
+\pagebreak
+
+\subsection{Experiments on Biological Data}
 

### 2012-12-20 ParetoHMM/HMM.py
Initial commits to code base. Defined skeletons for HMM,CMRF and Simulation outline

@@ -1,0 +1,67 @@
+""" 
+HMM, CMRF and TMRF classes are defined here
+
+@author: smoitra@cs.cmu.edu
+@license: BSD
+
+"""
+
+import os,sys
+import pylab as pl
+import numpy as np
+
+
+class HMM(object) : 
+	""" HMM """
+	def __init__(self) : 
+		""" Initialized using the train method """
+		self.trained = False
+
+	def train(self,traindata) : 
+	""" Train the HMM and set the params of the model"""
+		pass
+
+	def score(self,seq,obs) : 
+		""" Score a given sequence and observation"""
+		pass
+
+	def decode(self,obs)  : 
+		""" Perform max decoding to get the optimal sequence """
+		pass
+
+	def sample(self) :
+		""" Sample from the HMM"""
+		if not self.trained : 
+			sys.stderr.write("HMM not yet trained. Cannot Sample!")
+			return
+		pass
+
+class CMRF(HMM) : 
+	""" CMRF """
+	def __init__(self,hmm) : 
+		""" Converts hmm to CMRF """
+		pass
+
+	def train(self,traindata) : 
+	""" Train the HMM and set the params of the model"""
+		pass
+
+	def score(self,seq,obs) : 
+		""" Score a given sequence and observation"""
+		pass
+
+	def decode(self,obs)  : 
+		""" Perform max decoding to get the optimal sequence """
+		pass
+
+	def sample(self) :
+		""" Sample from the HMM"""
+		if not self.trained : 
+			sys.stderr.write("HMM not yet trained. Cannot Sample!")
+			return
+		pass
+
+
+
+
+

### 2012-12-20 ParetoHMM/sim1.py

@@ -1,0 +1,18 @@
+"""
+Simulation Study 1 :
+Create a 12 length protein
+
+@author: smoitra@cs.cmu.edu
+@license: BSD
+"""
+import os,sys
+import pylab as pl
+import numpy as np
+import scipt as sci
+
+
+
+if __name__ == '__main__' : 
+	pass
+
+

### 2012-12-20 ParetoHMM/HMM.py
Scoring function for the HMM

@@ -9,21 +9,45 @@ HMM, CMRF and TMRF classes are defined here
 import os,sys
 import pylab as pl
 import numpy as np
-
+import random
 
 class HMM(object) : 
 	""" HMM """
 	def __init__(self) : 
 		""" Initialized using the train method """
 		self.trained = False
+		self.length = None
+		self.emit = None # Emission probs
+		self.trans = None # Transition probs
+		self.dims = None # (latent,emit) dimspace
+		self.featmap = None # Maps features to ids
+		self.seqmap = None # Maps aa types to ids
+		self.initprob = None
+
 
 	def train(self,traindata) : 
-	""" Train the HMM and set the params of the model"""
+		""" Train the HMM and set the params of the model"""
 		pass
 
 	def score(self,seq,obs) : 
 		""" Score a given sequence and observation"""
-		pass
+		# Convert seq to ids
+		sids = [seqmap[aa] for seqmap,aa in zip(self.seqmap,seq)]	
+		# convert feature to ids
+		fids = [featmap[f] for featmap,f in zip(self.featmap,obs)]
+
+		ll = 0.0
+		# sum up scores of emission variables
+		for i in xrange(self.length) : 
+			ll += np.log(self.emit[i][sids[i]][fids[i]])
+		
+		# sum up score of latent variables
+		ll += np.log(self.initprob[sids[0]])
+		for i in xrange(i,self.length-1) : 
+			ll+= np.log(self.trans[i][sids[i]][sids[i+1]])
+
+		return ll
+
 
 	def decode(self,obs)  : 
 		""" Perform max decoding to get the optimal sequence """
@@ -31,10 +55,11 @@ class HMM(object) :
 
 	def sample(self) :
 		""" Sample from the HMM"""
-		if not self.trained : 
+		if not self.trained  : 
 			sys.stderr.write("HMM not yet trained. Cannot Sample!")
 			return
 		pass
+
 
 class CMRF(HMM) : 
 	""" CMRF """
@@ -43,7 +68,7 @@ class CMRF(HMM) :
 		pass
 
 	def train(self,traindata) : 
-	""" Train the HMM and set the params of the model"""
+		""" Train the HMM and set the params of the model"""
 		pass
 
[…]

### 2012-12-21 ParetoHMM/HMM.py
Added decoding skeleton

@@ -17,7 +17,7 @@ class HMM(object) :
 		""" Initialized using the train method """
 		self.trained = False
 		self.length = None
-		self.emit = None # Emission probs
+		self.emit = None # Emission probs (seq,feat)
 		self.trans = None # Transition probs
 		self.dims = None # (latent,emit) dimspace
 		self.featmap = None # Maps features to ids
@@ -51,7 +51,29 @@ class HMM(object) :
 
 	def decode(self,obs)  : 
 		""" Perform max decoding to get the optimal sequence """
-		pass
+		# convert feature to ids
+		fids = [featmap[f] for featmap,f in zip(self.featmap,obs)]
+		V = []
+		Ptr = []
+		# For the first element
+		V.append(map(lambda(k):np.log(self.emit[0][k][fids[0]])\
+			+np.log(self.initprob[k]),range(self.dims[0][0])))
+
+		# For the rest of the elements
+		for i in range(1,hmm.length) :
+			ptr_k = []; V_k = [];
+			for k in range(self.dims[i][0]) :
+				X = map(lambda(j):np.log(self.trans[i-1][j][k])+V[i-1][j],\
+					range(self.dims[i-1][0]))
+				maxX = max(X)
+				ptr_k.append(X.index(maxX))
+				V_k.append(np.log(self.emit[i][k][fids[i]])+maxX)
+			Ptr.append(Ptr)
+			V.append(V_k)
+
+		1/0
+
+
 
 	def sample(self) :
 		""" Sample from the HMM"""
@@ -142,6 +164,13 @@ if __name__ == '__main__' :
 		score = hmm.score(seq,feat)
 		print('Score for seq:{0} with feat:{1} is {2}'.\
 			format(seq,feat,score))
+
+
+	print("*"*10+"Decoding HMM"+"*"*10)
+	hmm.decode(feat1)	
+
+
+
 
 
 

### 2012-12-22 ParetoHMM/HMM.py
Started pointer retracing

@@ -71,8 +71,11 @@ class HMM(object) :
 			Ptr.append(Ptr)
 			V.append(V_k)
 
-		1/0
+		# retreive the optimal sequence
+		seq_max = []
+		seq_max.append(V[-1].index(max(V[-1])))
 
+		1/0
 
 
 	def sample(self) :
@@ -154,6 +157,9 @@ if __name__ == '__main__' :
 	cmrf = CMRF(hmm)
 	seq1 = 'a'*12
 	feat1 = 'HHHHLLLLHHHH'
+	seq2 = 'b'*12
+	feat2 = 'BBBBLLLLBBBB'
+
 	score1 = hmm.score(seq1,feat1)
 	print('Score for seq:{0} with feat:{1} is {2}'.format(seq1,feat1,score1))
 

### 2012-12-22 ParetoHMM/HMM.py
Finished decoding section. Bugfixed scoring

@@ -43,7 +43,7 @@ class HMM(object) :
 		
 		# sum up score of latent variables
 		ll += np.log(self.initprob[sids[0]])
-		for i in xrange(i,self.length-1) : 
+		for i in xrange(self.length-1) : 
 			ll+= np.log(self.trans[i][sids[i]][sids[i+1]])
 
 		return ll
@@ -68,15 +68,22 @@ class HMM(object) :
 				maxX = max(X)
 				ptr_k.append(X.index(maxX))
 				V_k.append(np.log(self.emit[i][k][fids[i]])+maxX)
-			Ptr.append(Ptr)
+			Ptr.append(ptr_k)
 			V.append(V_k)
 
 		# retreive the optimal sequence
 		seq_max = []
 		seq_max.append(V[-1].index(max(V[-1])))
-
-		1/0
-
+		energy_max = V[-1][seq_max[-1]]
+		for ptr_k in reversed(Ptr) :
+			seq_max.append(ptr_k[seq_max[-1]])
+		seq_max.reverse()
+		
+		# map it back to the original sequence
+		decode_seq = "".join([self.seqmap2[i][seqid] \
+			for i,seqid in enumerate(seq_max)])
+	
+		return energy_max,decode_seq
 
 	def sample(self) :
 		""" Sample from the HMM"""
@@ -140,6 +147,7 @@ def set_params_hmm_exp1(hmm) :
 		[[0.7,0.3],[0.3,0.7]]
 	]*hmm.length
 	hmm.seqmap = [{'a':0,'b':1}]*hmm.length
+	hmm.seqmap2 = [{0:'a',1:'b'}]*hmm.length
 	hmm.featmap = [{'H':0,'B':1,'L':2}]*hmm.length
 	hmm.initprob = [0.5,0.5]
 	hmm.trained = True
@@ -173,8 +181,11 @@ if __name__ == '__main__' :
 
 
 	print("*"*10+"Decoding HMM"+"*"*10)
-	hmm.decode(feat1)	
-
+	score,seq = hmm.decode(feat1)	
+	print("Decoded seq:{} energy:{} feat:{}".format(seq,score,feat1))
+	feat3 = "HHHHLLLLBBBB"
+	score,seq = hmm.decode(feat3)	
+	print("Decoded seq:{} energy:{} feat:{}".format(seq,score,feat3))
 
 
 

### 2012-12-23 ParetoHMM/HMM.py
Added some terms to CMRF

@@ -93,11 +93,17 @@ class HMM(object) :
 		pass
 
 
-class CMRF(HMM) : 
+class CMRF(object) : 
 	""" CMRF """
 	def __init__(self,hmm) : 
 		""" Converts hmm to CMRF """
-		pass
+		self.trained = hmm.trained
+		self.length = hmm.length
+		self.emit = hmm.emit # Emission probs (seq,feat)
+		self.trans = hmm.trans # Transition probs
+		self.dims = hmm.dims # (latent,emit) dimspace
+		self.featmap = hmm.featmap # Maps features to ids
+		self.seqmap = hmm.seqmap # Maps aa types to ids
 
 	def train(self,traindata) : 
 		""" Train the HMM and set the params of the model"""

### 2012-12-23 ParetoHMM/HMM.py
Added scoring to CMRF

@@ -99,19 +99,73 @@ class CMRF(object) :
 		""" Converts hmm to CMRF """
 		self.trained = hmm.trained
 		self.length = hmm.length
-		self.emit = hmm.emit # Emission probs (seq,feat)
-		self.trans = hmm.trans # Transition probs
+		self.emitpsi = hmm.emit # Emission probs (seq,feat)
+		self.seqpsi = hmm.trans # Transition probs
 		self.dims = hmm.dims # (latent,emit) dimspace
 		self.featmap = hmm.featmap # Maps features to ids
 		self.seqmap = hmm.seqmap # Maps aa types to ids
-
-	def train(self,traindata) : 
-		""" Train the HMM and set the params of the model"""
-		pass
+		self.seqphi = [np.ones(dim[0]) for dim in hmm.dims]
+		self.featphi =[np.ones(dim[1]) for dim in hmm.dims] 
 
 	def score(self,seq,obs) : 
 		""" Score a given sequence and observation"""
-		pass
+		# Convert seq to ids
+		sids = [seqmap[aa] for seqmap,aa in zip(self.seqmap,seq)]	
+		# convert feature to ids
+		fids = [featmap[f] for featmap,f in zip(self.featmap,obs)]
+
+		ll = 0.0
+		# sum up scores of emission variables
+		for i in xrange(self.length) : 
+			ll += np.log(self.emitpsi[i][sids[i]][fids[i]])
+		
+		# sum up score of latent variables
+		ll += np.log(0.5)
+		for i in xrange(self.length-1) : 
+			ll+= np.log(self.seqpsi[i][sids[i]][sids[i+1]])
+
+		# sum up the scores of the nodes
+		for i in xrange(self.length) : 
+			ll+=  np.log(self.seqphi[i][sids[i]])
+			ll+= np.log(self.featphi[i][fids[i]])
+		return ll
+
+
+	def decode(self,obs)  : 
+		""" Perform max decoding to get the optimal sequence """
+		# convert feature to ids
+		fids = [featmap[f] for featmap,f in zip(self.featmap,obs)]
+		V = []
+		Ptr = []
+		# For the first element
+		V.append(map(lambda(k):np.log(self.emit[0][k][fids[0]])\
+			+np.log(self.initprob[k]),range(self.dims[0][0])))
+
+		# For the rest of the elements
[…]

### 2012-12-23 ParetoHMM/HMM.py
CMRF decoding code finished. Added TMRF skeleton

@@ -11,6 +11,11 @@ import pylab as pl
 import numpy as np
 import random
 
+
+def reverse_dict(d)	:	
+	""" Reverses a dictionary making the keys be the values"""
+	return dict([(v,k) for k,v in d.viewitems()])	
+
 class HMM(object) : 
 	""" HMM """
 	def __init__(self) : 
@@ -23,7 +28,8 @@ class HMM(object) :
 		self.featmap = None # Maps features to ids
 		self.seqmap = None # Maps aa types to ids
 		self.initprob = None
-
+		self.seqmap2 = None
+		self.fratmap2 = None
 
 	def train(self,traindata) : 
 		""" Train the HMM and set the params of the model"""
@@ -63,7 +69,8 @@ class HMM(object) :
 		for i in range(1,hmm.length) :
 			ptr_k = []; V_k = [];
 			for k in range(self.dims[i][0]) :
-				X = map(lambda(j):np.log(self.trans[i-1][j][k])+V[i-1][j],\
+				X = map(lambda(j):\
+					np.log(self.trans[i-1][j][k])+V[i-1][j],\
 					range(self.dims[i-1][0]))
 				maxX = max(X)
 				ptr_k.append(X.index(maxX))
@@ -92,7 +99,6 @@ class HMM(object) :
 			return
 		pass
 
-
 class CMRF(object) : 
 	""" CMRF """
 	def __init__(self,hmm) : 
@@ -106,6 +112,10 @@ class CMRF(object) :
 		self.seqmap = hmm.seqmap # Maps aa types to ids
 		self.seqphi = [np.ones(dim[0]) for dim in hmm.dims]
 		self.featphi =[np.ones(dim[1]) for dim in hmm.dims] 
+
+		# maps from id to seq
+		self.seqmap2 = map(reverse_dict,self.seqmap)
+		self.featmap2 = map(reverse_dict,self.featmap)
 
 	def score(self,seq,obs) : 
 		""" Score a given sequence and observation"""
@@ -138,18 +148,18 @@ class CMRF(object) :
 		V = []
 		Ptr = []
 		# For the first element
-		V.append(map(lambda(k):np.log(self.emit[0][k][fids[0]])\
-			+np.log(self.initprob[k]),range(self.dims[0][0])))
+		V.append(map(lambda(k):np.log(self.emitpsi[0][k][fids[0]])\
+			+np.log(self.seqphi[k]),range(self.dims[0][0])))
 
[…]

### 2012-12-23 docs/tex/draft.tex
Added scoring and decoding for TMRF. Fixed scoring and decoding for CMRF

@@
 
+\appendix
+
+\section{Other Notes}
+
+\subsection{Implementation Notes}
+
+\paragraph{Implementation Scheme 1} : Right now there is a single $\phi,\psi$ parameter set.
+This means that you use a single alignment for learning the parameters of the HMM.
+Once the parameter have been learned the energy functions are defined by conditioning on the features of the state you are interested in.
+The parameters of the different amino acids are tied together.
+This need not be always be the case.
+
+Instead of having different nodes for all the the different TMRFs.
+The energy functions can be so modified such that they correspond to a single CMRF.
+then decoding and sampling from this CMRF is the same as before.
+
+This can be done as 
+\begin{align*}
+\phi^{TMRF}(X) &= 1 \\
+\phi^{TMRF}(\mathbf{O_i}) &= 1 \\
+\psi^{TMRF}(X_i,X_{i+1}) &= \psi(X_i,X_j)\\
+\psi^{TMRF}(X_i,\mathbf{O_i^{AB}}) &= \psi(X_i,\mathbf{O_i^A})^{\theta_A}\psi(X_i,\mathbf{O_i^B})^{\theta_B}
+\end{align*}
+
+\paragraph{Implementation Scheme 2} : Different alignments that represent different amino acid factors for the protein class under consideration.
+This gives rise to different $\phi,\psi$ features.
+How does this change inference and decoding ?
+
+\paragraph{Implementation Scheme 3} : Different node and edge factors for the different amoino acids. This means that the factors are not tied together.
+
+\paragraph{Decode All seq} : It may be possible to find all the sequences that give the same energy rather than just return only one of them.
+this can be done by storing a list of pointers during the DP decoding.
+
+\subsection{Experiment Design Notes}
+
+\paragraph{Benchmarking} : It is important to compare the performance of our algorithm with some other well known protein design algorithms.
+This is important from the purpose of getting a paper accepted.
+The more evidence you can provide the more convincing a paper is.
+
 \end{document}

### 2012-12-23 ParetoHMM/HMM.py

@@ -130,7 +130,6 @@ class CMRF(object) :
 			ll += np.log(self.emitpsi[i][sids[i]][fids[i]])
 		
 		# sum up score of latent variables
-		ll += np.log(0.5)
 		for i in xrange(self.length-1) : 
 			ll+= np.log(self.seqpsi[i][sids[i]][sids[i+1]])
 
@@ -149,7 +148,7 @@ class CMRF(object) :
 		Ptr = []
 		# For the first element
 		V.append(map(lambda(k):np.log(self.emitpsi[0][k][fids[0]])\
-			+np.log(self.seqphi[k]),range(self.dims[0][0])))
+			+np.log(self.seqphi[0][k]),range(self.dims[0][0])))
 
 		# For the rest of the elements
 		for i in range(1,hmm.length) :
@@ -159,7 +158,8 @@ class CMRF(object) :
 					range(self.dims[i-1][0]))
 				maxX = max(X)
 				ptr_k.append(X.index(maxX))
-				V_k.append(np.log(self.emitpsi[i][k][fids[i]])+maxX)
+				V_k.append(np.log(self.seqphi[i][k])+\
+					np.log(self.emitpsi[i][k][fids[i]])+maxX)
 			Ptr.append(ptr_k)
 			V.append(V_k)
 
@@ -186,36 +186,42 @@ class CMRF(object) :
 
 class TMRF(object) : 
 	""" The Tree MRF which can be used for decoding """
-	def __init__(self,*cmrflist) : 
+	def __init__(self,cmrf,thetalist,featlist) : 
 		""" Converts A list of cmrfs to a joint TMRF """
-		self.trained = hmm.trained
-		self.length = hmm.length
-		self.emitpsi = hmm.emit # Emission probs (seq,feat)
-		self.seqpsi = hmm.trans # Transition probs
-		self.dims = hmm.dims # (latent,emit) dimspace
-		self.featmap = hmm.featmap # Maps features to ids
-		self.seqmap = hmm.seqmap # Maps aa types to ids
-		self.seqphi = [np.ones(dim[0]) for dim in hmm.dims]
-		self.featphi =[np.ones(dim[1]) for dim in hmm.dims] 
+		self.nStates = len(thetalist)
+		assert(self.nStates>1)
[…]

### 2012-12-23 docs/tex/draft.tex
Added some comments about theoretical details

@@
 \paragraph{Decode All seq} : It may be possible to find all the sequences that give the same energy rather than just return only one of them.
-this can be done by storing a list of pointers during the DP decoding.
+this can be done by storing a list of pointers during the DP
+decoding.
+
+\subsection{Theory Notes}
+
+\paragraph{Failure Cases} : Here we list some of the failure cases : 
+\begin{itemize}
+\item Can you prove that the number of points on the convex frontier are not exponential ?
+\item This algorithm is slow when the points on the frontier resemble the circumeference of a circle
+\end{itemize}
+
+\paragraph{More than 2 states } : how does the convex hull algo work for more than 2 states ? How do you calculate the slope of a 3 dimensional polytope ?
+
+\paragraph{convex combinations?} : Does $\theta_A$ and $\theta_B$ need to sum to one?
 

### 2012-12-23 ParetoHMM/sim1.py

@@ -8,11 +8,53 @@ Create a 12 length protein
 import os,sys
 import pylab as pl
 import numpy as np
-import scipt as sci
+import scipy as sci
+from HMM import HMM,CMRF,TMRF
+from itertools import product
 
-
+def set_params_hmm_exp1(hmm) : 
+	""" Sets the params of a hmm for sim experiment 1"""
+	hmm.length = 12
+	hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
+	hmm.emit = [
+		[[0.6,0.2,0.2],[0.2,0.6,0.2]]
+	]*hmm.length
+	hmm.trans = [
+		[[0.7,0.3],[0.3,0.7]]
+	]*hmm.length
+	hmm.seqmap = [{'a':0,'b':1}]*hmm.length
+	hmm.seqmap2 = [{0:'a',1:'b'}]*hmm.length
+	hmm.featmap = [{'H':0,'B':1,'L':2}]*hmm.length
+	hmm.initprob = [0.5,0.5]
+	hmm.trained = True
 
 if __name__ == '__main__' : 
-	pass
+	hmm = HMM()
+	# Set the params of the h,,
+	set_params_hmm_exp1(hmm)
+	cmrf = CMRF(hmm)
+	seq1 = 'a'*12
+	feat1 = 'HHHHLLLLHHHH'
+	seq2 = 'b'*12
+	feat2 = 'BBBBLLLLBBBB'
 
 
+	
+	
+	# Plot the entire sequence space
+	ll_list1,ll_list2 = [],[]
+	for seq in product('ab',repeat=12):	
+		ll_list1.append(cmrf.score(seq,feat1))
+		ll_list2.append(cmrf.score(seq,feat2))
+
+	pl.figure()
+	pl.plot(ll_list1,ll_list2,'b*')
+	pl.xlabel('Energy:'+feat1)
+	pl.ylabel('Energy:'+feat2)
+	pl.title('Energy Plot')
+	xmin,xmax = pl.xlim()
+	ymin,ymax = pl.ylim()
+	pl.xlim(xmin,2)
+	pl.ylim(ymin,2)
+	pl.axvline()
+	pl.axhline()

### 2012-12-24 ParetoHMM/HMM.py
Pareto frontier code

@@ -137,7 +137,7 @@ class CMRF(object) :
 		for i in xrange(self.length) : 
 			ll+=  np.log(self.seqphi[i][sids[i]])
 			ll+= np.log(self.featphi[i][fids[i]])
-		return ll
+		return -ll
 
 
 	def decode(self,obs)  : 
@@ -151,7 +151,7 @@ class CMRF(object) :
 			+np.log(self.seqphi[0][k]),range(self.dims[0][0])))
 
 		# For the rest of the elements
-		for i in range(1,hmm.length) :
+		for i in range(1,self.length) :
 			ptr_k = []; V_k = [];
 			for k in range(self.dims[i][0]) :
 				X = map(lambda(j):np.log(self.seqpsi[i-1][j][k])+V[i-1][j],\
@@ -175,7 +175,7 @@ class CMRF(object) :
 		decode_seq = "".join([self.seqmap2[i][seqid] \
 			for i,seqid in enumerate(seq_max)])
 	
-		return energy_max,decode_seq
+		return -energy_max,decode_seq
 
 	def sample(self) :
 		""" Sample from the HMM"""
@@ -229,7 +229,7 @@ class TMRF(object) :
 		for i in xrange(self.length) : 
 			ll+=  np.log(self.seqphi[i][sids[i]])
 			ll+= np.log(self.featphi[i][fids[i]])
-		return ll
+		return -ll
 		
 
 	def decode(self)  : 
@@ -253,7 +253,7 @@ class TMRF(object) :
 		V.append(v_k)
 
 		# For the rest of the elements
-		for i in range(1,hmm.length) :
+		for i in range(1,self.length) :
 			ptr_k = []; V_k = [];
 			for k in range(self.dims[i][0]) :
 				X = map(lambda(j):np.log(self.seqpsi[i-1][j][k])\
@@ -278,7 +278,7 @@ class TMRF(object) :
 		decode_seq = "".join([self.seqmap2[i][seqid] \
 			for i,seqid in enumerate(seq_max)])
 	
-		return energy_max,decode_seq
+		return -energy_max,decode_seq
 
 	def sample(self) :
 		""" Sample from the HMM"""

### 2012-12-24 ParetoHMM/cvxhull.py

@@ -1,0 +1,90 @@
+"""
+Contains all the routines for finding the convex hull/ pareto frontier
+
+@author:smoitra@cs.cmu.edu
+"""
+
+import numpy as np
+import os,sys
+import pylab as pl
+from HMM import HMM,CMRF, TMRF
+from pdb import set_trace as stop
+
+
+def pareto_frontier(cmrf,featlist) :
+	"""Finds and prints the pareto frontier. Currently works for 
+		two competing states only
+	 """
+	Q = []
+	nStates = len(featlist)
+	feat1,feat2 = featlist
+	Eaxa,Xa = cmrf.decode(feat1)
+	Ebxb,Xb = cmrf.decode(feat2)
+	Eaxb = cmrf.score(Xa,feat2)
+	Ebxa = cmrf.score(Xb,feat1)
+	Q.append((Xa,Xb))
+	frontier,frontier_energy = [],[]
+	frontier.extend([Xa,Xb])
+	frontier_energy.extend([(Eaxa,Ebxa),(Eaxb,Ebxb)])
+	while len(Q) > 0 :
+		### Optimize 
+		Xa,Xb = Q[0]
+		Q = Q[1:] # Dequeue
+		Eaxb = cmrf.score(Xb,feat1)
+		Ebxa = cmrf.score(Xa,feat2)	
+		Eaxa = cmrf.score(Xa,feat1)
+		Ebxb = cmrf.score(Xb,feat2)	
+
+		m = (Ebxa - Ebxb)/(Eaxa-Eaxb)
+		thetaa = -m/(1-m)
+		thetab = 1/(1-m)
+		tmrf = TMRF(cmrf,[thetaa,thetab],[feat1,feat2])
+		Xab = tmrf.decode()[1]
+		if Xab != Xa and Xab != Xb : 
+			frontier.append(Xab)
+			Eaxab = cmrf.score(Xab,feat1)
+			Ebxab = cmrf.score(Xab,feat2)
+			frontier_energy.append((Eaxab,Ebxab))
+			Q.extend([(Xa,Xab),(Xab,Xb)])
+	# Calculate energy of frontier elements	
+	return frontier,frontier_energy
+
+
+def set_params_hmm_exp1(hmm) : 
+	""" Sets the params of a hmm for sim experiment 1"""
+	hmm.length = 12
[…]

### 2012-12-24 ParetoHMM/sim1.py

@@ -11,6 +11,7 @@ import numpy as np
 import scipy as sci
 from HMM import HMM,CMRF,TMRF
 from itertools import product
+from cvxhull import pareto_frontier 
 
 def set_params_hmm_exp1(hmm) : 
 	""" Sets the params of a hmm for sim experiment 1"""
@@ -38,23 +39,26 @@ if __name__ == '__main__' :
 	seq2 = 'b'*12
 	feat2 = 'BBBBLLLLBBBB'
 
-
-	
-	
 	# Plot the entire sequence space
 	ll_list1,ll_list2 = [],[]
 	for seq in product('ab',repeat=12):	
 		ll_list1.append(cmrf.score(seq,feat1))
 		ll_list2.append(cmrf.score(seq,feat2))
 
+	# Find the pareto frontier
+	frontier,frontier_energy = pareto_frontier(cmrf,[feat1,feat2])
+
 	pl.figure()
 	pl.plot(ll_list1,ll_list2,'b*')
+	pl.plot(*list(zip(*frontier_energy)),color='magenta',linestyle='*')
+	
 	pl.xlabel('Energy:'+feat1)
 	pl.ylabel('Energy:'+feat2)
 	pl.title('Energy Plot')
 	xmin,xmax = pl.xlim()
 	ymin,ymax = pl.ylim()
-	pl.xlim(xmin,2)
-	pl.ylim(ymin,2)
+	pl.xlim(-2,xmax)
+	pl.ylim(-2,ymax)
 	pl.axvline()
 	pl.axhline()
+	

### 2012-12-24 ParetoHMM/sim2.py
Adding second simulation with random prob values. results unstable

@@ -1,5 +1,5 @@
 """
-Simulation Study 1 :
+Simulation Study 2 :
 Create a 12 length protein
 
 @author: smoitra@cs.cmu.edu
@@ -12,16 +12,23 @@ import scipy as sci
 from HMM import HMM,CMRF,TMRF
 from itertools import product
 from cvxhull import pareto_frontier 
+import random
+
+
+def gen_random_dist(size) : 
+	x = np.random.uniform(size=size)
+	x = x/sum(x)
+	return x
 
 def set_params_hmm_exp1(hmm) : 
 	""" Sets the params of a hmm for sim experiment 1"""
 	hmm.length = 12
 	hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
 	hmm.emit = [
-		[[0.6,0.2,0.2],[0.2,0.6,0.2]]
+		[gen_random_dist(3),gen_random_dist(3)]
 	]*hmm.length
 	hmm.trans = [
-		[[0.7,0.3],[0.3,0.7]]
+		[gen_random_dist(2),gen_random_dist(2)]
 	]*hmm.length
 	hmm.seqmap = [{'a':0,'b':1}]*hmm.length
 	hmm.seqmap2 = [{0:'a',1:'b'}]*hmm.length
@@ -62,4 +69,5 @@ if __name__ == '__main__' :
 	pl.ylim(-2,ymax)
 	pl.axvline()
 	pl.axhline()
-	pl.savefig('../docs/tex/pics/sim1.pdf')
+	pl.savefig('../docs/tex/pics/sim2.pdf')
+	pl.show()

### 2012-12-27 docs/tex/draft.tex
Pareto frontier code
Debugging a case wherein multiple sequences with the same energy cause the CMRF decoding to fail

@@
 \subsubsection{Converting a HMM to CMRF}
-It is possible to reparameterize an HMM so that it now becomes a CMRF. The partition function $Z$ of the newly constructed CMRF equals 1 corresponding to a probability distribution., However there is
+It is possible to reparameterize an HMM so that it now becomes a CMRF. The partition function $Z$ of the newly constructed CMRF equals 1 corresponding to a probability distribution.
 
-\subsection{Combining the CMRF to a HMM}
+\subsection{Combining CMRFs to get a TMRF}
 
@@
 
+\paragraph{Multiple sequence Same Energy} :  Currently the cvhull algorithm is being thrown out of gear by sequences that have the same energy.
+An example of this situation is shown in Figure~\ref{fig:sim2_debug1}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=1.0\textwidth]{pics/sim2_debug1.pdf}
+    \caption{This figure illustrates the issue when there are multiple sequences with the same energy.
+Here sequences 3 and 4 have the same energy.
+This causes the convex hull algorithm to fail in its current implementation.
+}
+    \label{fig:sim2_debug1}
+\end{figure}
+\pagebreak
+
 \subsection{Theory Notes}

### 2012-12-27 ParetoHMM/cvxhull.py

@@ -9,7 +9,7 @@ import os,sys
 import pylab as pl
 from HMM import HMM,CMRF, TMRF
 from pdb import set_trace as stop
-
+import warnings
 
 def pareto_frontier(cmrf,featlist) :
 	"""Finds and prints the pareto frontier. Currently works for 
@@ -20,8 +20,10 @@ def pareto_frontier(cmrf,featlist) :
 	feat1,feat2 = featlist
 	Eaxa,Xa = cmrf.decode(feat1)
 	Ebxb,Xb = cmrf.decode(feat2)
-	Eaxb = cmrf.score(Xa,feat2)
-	Ebxa = cmrf.score(Xb,feat1)
+	if Xa == Xb : 
+		return [Xa],[(Eaxa,Ebxb)]
+	Eaxb = cmrf.score(Xb,feat1)
+	Ebxa = cmrf.score(Xa,feat2)
 	Q.append((Xa,Xb))
 	frontier,frontier_energy = [],[]
 	frontier.extend([Xa,Xb])
@@ -34,8 +36,11 @@ def pareto_frontier(cmrf,featlist) :
 		Ebxa = cmrf.score(Xa,feat2)	
 		Eaxa = cmrf.score(Xa,feat1)
 		Ebxb = cmrf.score(Xb,feat2)	
-
 		m = (Ebxa - Ebxb)/(Eaxa-Eaxb)
+		if m > 0 : 
+			#stop()
+			sys.stderr.write("### WARNING : Slope > 0. Cvxhull failed")
+			return frontier,frontier_energy
 		thetaa = -m/(1-m)
 		thetab = 1/(1-m)
 		tmrf = TMRF(cmrf,[thetaa,thetab],[feat1,feat2])

### 2012-12-27 ParetoHMM/sim2.py

@@ -1,6 +1,7 @@
 """
 Simulation Study 2 :
 Create a 12 length protein
+Run Pareto Frontier
 
 @author: smoitra@cs.cmu.edu
 @license: BSD
@@ -13,7 +14,7 @@ from HMM import HMM,CMRF,TMRF
 from itertools import product
 from cvxhull import pareto_frontier 
 import random
-
+DEBUG=1
 
 def gen_random_dist(size) : 
 	x = np.random.uniform(size=size)
@@ -46,28 +47,52 @@ if __name__ == '__main__' :
 	seq2 = 'b'*12
 	feat2 = 'BBBBLLLLBBBB'
 
+	### DEBUG
+	import pickle
+	cmrf = pickle.load(open('cmrf.pkl'))
+
+
 	# Plot the entire sequence space
 	ll_list1,ll_list2 = [],[]
-	for seq in product('ab',repeat=12):	
+	seq_list = ["".join(s) for s in product('ab',repeat=12)]
+	for seq in seq_list:	
 		ll_list1.append(cmrf.score(seq,feat1))
 		ll_list2.append(cmrf.score(seq,feat2))
+
+	min_feat1id = ll_list1.index(min(ll_list1))
+	min_feat2id = ll_list2.index(min(ll_list2))
 
 	# Find the pareto frontier
 	frontier,frontier_energy = pareto_frontier(cmrf,[feat1,feat2])
 
 	pl.figure()
-	pl.plot(ll_list1,ll_list2,'b*')
+	#pl.plot(ll_list1,ll_list2,'b*')
 	pl.plot(*zip(*frontier_energy),color='magenta',marker='*',\
 		linestyle='dashed')
-	
+	if DEBUG :
+		ctr = dict(zip(set(frontier_energy),[0]*\
+			len(set(frontier_energy))))
+		for i,e in enumerate(frontier_energy) : 
+			ctr[e] += 1
+			pl.text(e[0],e[1]+0.1*ctr[e],str(i),fontsize=8)
+			pl.text(e[0]+0.1,e[1]+0.1*ctr[e],frontier[i],fontsize=7)	
+
 	pl.xlabel('Energy:'+feat1)
 	pl.ylabel('Energy:'+feat2)
 	pl.title('Energy Plot')
 	xmin,xmax = pl.xlim()
 	ymin,ymax = pl.ylim()
[…]

### 2012-12-27 docs/tex/draft.tex
Finished Simulation 1, sim 2 debugging and results, sim3 skeleton. Added changes to document

@@
 \subsubsection{Training an HMM}
+Training an HMM requires the following 
+\begin{itemize}
+\item Protein sequence alignments
+\item A mapping between amino acids of the sequences to discrete features
+\end{itemize}
+
+The emission and transition probabilities are calculated as
+
+\begin{align*}
+P(X_{i+1}=b|X_i=a) = \frac{\text{count}(X_{i+1}=b)+d}{\text{count}(X_i=a)+dS} \\
+P(O_{i}=f|X_i=a) = \frac{\text{count}(O_{i}=f)+d}{\text{count}(X_i=a)+dF}
+\end{align*}
+where d is a smoothing count and $S$ and $F$ are the sizes of the alphabet of the sequence space and the feature space respectively.
 
@@
 It is possible to reparameterize an HMM so that it now becomes a CMRF. The partition function $Z$ of the newly constructed CMRF equals 1 corresponding to a probability distribution.
+\\
+\\
+This can be done as :
+\begin{align*}
+\phi(X) &= 1 \\
+\phi(\mathbf{O_i})&=1 \\
+\psi(X_i,X_{i+1}) &= P(X_i,X_j)\\
+\psi(X_i,\mathbf{O_i}) &= P(\mathbf{O_i}|X_i)
+\end{align*}
 
 \subsection{Combining CMRFs to get a TMRF}
+Instead of having different nodes for all the the different TMRFs.
+The energy functions can be so modified such that they correspond to a single CMRF.
+then decoding and sampling from this CMRF is the same as before.
+\\
+\\
+This can be done as 
+\begin{align*}
+\phi^{TMRF}(X) &= 1 \\
+\phi^{TMRF}(\mathbf{O_i}) &= 1 \\
+\psi^{TMRF}(X_i,X_{i+1}) &= \psi(X_i,X_j)\\
+\psi^{TMRF}(X_i,\mathbf{O_i^{AB}}) &= \psi(X_i,\mathbf{O_i^A})^{\theta_A}\psi(X_i,\mathbf{O_i^B})^{\theta_B}
+\end{align*}
+
+\subsection{Augmenting Feature Space}
+The feature space can be augmented to include any set of discrete features $F_1,F_2,\cdots,F_k$.
+Such a feature space can be added on to the existing feature space using a tree-like structure as shown in Figure~\ref{fig:CMRF_augmented}.
+\\
+\\
+The probability of a sequence assignment is defined as follows :
+\begin{align*}
+P(X_1,X_2,\dotsc,X_n) &= \frac{1}{Z}\prod_{i=1}^{n} \phi(X_i)\left(\prod_{j=1}^{k}\phi(\mathbf{O_{iF_j}})\psi(\mathbf{O_{iF_j}},X_i)\right) \prod_{i=1}^{n-1}{\psi(X_i,X_{i+1})} 
+\end{align*}
+
+\begin{figure}[h]
+% selecting unit length
+\centering
+% used for centering Figure
+\setlength{\unitlength}{0.3cm}
+\begin{picture}(40,20)(0,0)
+\multiput(7,15)(8,0){1}{\circle{4}}
+\multiput(3,7)(8,0){1}{\circle{4}}
+\multiput(3,7)(8,0){1}{\circle{3.7}}
+\multiput(11,7)(8,0){1}{\circle{4}}
+\multiput(11,7)(8,0){1}{\circle{3.7}}
+\thicklines
+\multiput(9,15)(8,0){1}{\line(1,0){10}}
+\multiput(20,15)(1,0){4}{\circle*{0.3}}
+\multiput(7,13)(8,0){1}{\line(-1,-1){4}}
+\multiput(7,13)(8,0){1}{\line(0,-1){4}}
+\multiput(7,13)(8,0){1}{\line(1,-1){4}}
+\multiput(6,7)(1,0){3}{\circle*{0.3}}
+\put(24,15){\line(1,0){4}}
+\put(30,13){\line(-1,-1){4}}
+\put(30,13){\line(1,-1){4}}
+\put(30,13){\line(0,-1){4}}
+\multiput(29,7)(1,0){3}{\circle*{0.3}}
+\thinlines
+\put(30,15){\circle{4}}
+\put(26,7){\circle{4}}
+\put(26,7){\circle{3.7}}
+\put(34,7){\circle{4}}
+\put(34,7){\circle{3.7}}
+\put(5.5,14.6){$\phi(X_1)$}
+\put(28.5,14.6){$\phi(X_n)$}
+\put(0.5,6.6){$\phi(O_{1F_1})$}
+\put(8.5,6.6){$\phi(O_{1F_k})$}
+\put(23.5,6.6){$\phi(O_{nF_1})$}
+\put(31.5,6.6){$\phi(O_{nF_k})$}
+\put(22.5,16.5){\footnotesize{$\psi(X_{n{-}1}{,}X_n)$}}
+\put(-1,10.6){\footnotesize{$\psi(X_1{,}O_{1F_1})$}}
+\put(10,10.6){\footnotesize{$\psi(X_1{,}O_{1F_k})$}}
+\put(33.2,10.6){\footnotesize{$\psi(X_n{,}O_{nF_1})$}}
+\put(21.2,10.6){\footnotesize{$\psi(X_n{,}O_{nF_k})$}}
+\end{picture}
+\label{fig:CMRF_augmented}
+\caption{A CMRF with augmented feature space. Note the tree like shape of the feature space. Here $F_1,F_2,\dots,F_k$ correspond to different discrete features. }
+\end{figure}
+
+\pagebreak
 
@@
 \subsection{Simulation Experiments}
+We run a variety of different simulation experiments to check the performance of our algorithm on
+
+\subsubsection{Sim Experiment 1}
 We demonstrate how our algorithm performs on simulated data.
@@
 We also calculate the energies of all the $2^{12}$ sequences possible under this restricted sequence model and plot the points.
-The \texttt{Pareto-Frontier} algorithm was able to find the pareto frontier in \textcolor{red}{N steps} steps.
+The \texttt{Pareto-Frontier} algorithm was able to find the pareto frontier in \textcolor{red}{1 step} steps.
 See Figure~\ref{fig:pareto-sim} for the trace taken by the algorithm.
@@
     \centering
-    \includegraphics[width=0.9\textwidth]{pics/convexfront2.pdf}
-    \caption{ }
+    \includegraphics[width=0.9\textwidth]{pics/sim1.pdf}
+    \caption{Simulation Experiment 1: The emission and transition weights are as defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2}.
+The Pareto-HMM finds the optimal pareto frontier for a twelve length sequence using the alphabet \texttt\{a,b\} and the feature space \texttt\{H,L,B\} in two steps of the Convex hull algorithm.
+The magenta stars corresponds to the optimal sequences.
+The magenta line corresponds to the lower left convex hull of the sequence energy space.
+The blue stars correspond to non-optimal dominated sequences.
+The Pareto-Frontier algoerithm provably finds the frontier quickly instead of having to enumerate,calculate and sort the energies of a  $2^{12}$ sequence space.}
     \label{fig:pareto-sim}
 \end{figure}
+
 \pagebreak
+
+\subsubsection{Simulation Experiment 2}
+In this example we look at a slightly more complicated energy frontier.
+To generate this energy surface we resorted to generating random emission and transition probabilities till we found a sufficiently complicated energy frontier but still having alphabet \texttt{\{a,b\}} and feature space \texttt{\{H,L,B\}}.
+The probabilities for this simulation experiment are shown in tables~\ref{tab:psi3} and~\ref{tab:psi4}.
+
+\begin{table}
+\begin{center}
+\begin{tabular}{cc|c|c|c|}
+\cline{3-5}
+& & \multicolumn{3}{c|}{Observations($\mathbf{O_i}$)} \\ \cline{3-5}
+& & Helix(H) & Beta(B) & Loop(L) \\ \cline{1-5}
+\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
+\multicolumn{1}{|c|}{$\alpha$} & 0.334 & 0.272 & 0.394      \\ \cline{2-5}
+\multicolumn{1}{|c|}{}                        &
+\multicolumn{1}{|c|}{$\beta$} & 0.477 & 0.093 & 0.430      \\ \cline{1-5}
+\end{tabular}
+\end{center}
+\caption{$\psi(X,\mathbf{O})$ contains the weights for edge potentials of sequence and features together }
+\label{tab:psi3}
+\end{table}
+
+\begin{table}
+\begin{center}
+\begin{tabular}{cc|c|c|}
+\cline{3-4}
+& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
+& & $\alpha$ & $\beta$  \\ \cline{1-4}
+\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
+\multicolumn{1}{|c|}{$\alpha$} & 0.483 & 0.517      \\ \cline{2-4}
+\multicolumn{1}{|c|}{}                        &
+\multicolumn{1}{|c|}{$\beta$} & 0.589 & 0.411      \\ \cline{1-4}
+\end{tabular}
+\end{center}
+\caption{$\psi(X_i,X_{i+1})$ contains the weights for the edge potentials of adjacent amino acids}
+\label{tab:psi4}
+\end{table}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{pics/sim2_all.pdf}
+    \caption{Simulation Experiment 2: Pareto Frontier algorithm is applied to a more complicated energy surface.
+The algorithm finds the pareto frontier denoted by the dashed magenta in 5 steps, which is significantly faster than brute force enumeration.
+}
+    \label{fig:sim2_all}
+\end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{pics/sim2_frontier.pdf}
+    \caption{Simulation Experiment 2(Frontier): This shows the results from simulation 2. The sequences along the frontier are shown as well }
+    \label{fig:sim2_frontier}
+\end{figure}
+
+\pagebreak
+\clearpage
 
@@
 
+\subsubsection{Continuous Feature Spaces}
+Currently we are unable to provide support for continuous feature spaces.  Performing inference on discrete spaces by conditioning on continuous feature spaces can be more challenging.
+
 \section{Conclusion}
@@
 \paragraph{Multiple sequence Same Energy} :  Currently the cvhull algorithm is being thrown out of gear by sequences that have the same energy.
-An example of this situation is shown in Figure~\ref{fig:sim2_debug1}
+An example of this situation is shown in Figure~\ref{fig:sim2_debug1}.
+The current way we are dealing with this is to define a taboo list of energies that are to be ignored if they are found.
+A better way to do this would be to find all sequences with the same energy during the decoding stage and still use the same energy taboo list.
+
 \begin{figure}[h!]

### 2012-12-27 ParetoHMM/cvxhull.py

@@ -16,6 +16,7 @@ def pareto_frontier(cmrf,featlist) :
 		two competing states only
 	 """
 	Q = []
+	taboodict = {}
 	nStates = len(featlist)
 	feat1,feat2 = featlist
 	Eaxa,Xa = cmrf.decode(feat1)
@@ -28,6 +29,8 @@ def pareto_frontier(cmrf,featlist) :
 	frontier,frontier_energy = [],[]
 	frontier.extend([Xa,Xb])
 	frontier_energy.extend([(Eaxa,Ebxa),(Eaxb,Ebxb)])
+	taboodict[(Eaxa,Ebxa)] = 1;
+	taboodict[(Eaxb,Ebxb)] = 1;
 	while len(Q) > 0 :
 		### Optimize 
 		Xa,Xb = Q[0]
@@ -45,11 +48,13 @@ def pareto_frontier(cmrf,featlist) :
 		thetab = 1/(1-m)
 		tmrf = TMRF(cmrf,[thetaa,thetab],[feat1,feat2])
 		Xab = tmrf.decode()[1]
-		if Xab != Xa and Xab != Xb : 
+		Eaxab = cmrf.score(Xab,feat1)
+		Ebxab = cmrf.score(Xab,feat2)
+		if Xab != Xa and Xab != Xb and \
+			not taboodict.has_key((Eaxab,Ebxab)): 
 			frontier.append(Xab)
-			Eaxab = cmrf.score(Xab,feat1)
-			Ebxab = cmrf.score(Xab,feat2)
 			frontier_energy.append((Eaxab,Ebxab))
+			taboodict[(Eaxab,Ebxab)]=1
 			Q.extend([(Xa,Xab),(Xab,Xb)])
 	# Calculate energy of frontier elements	
 	return frontier,frontier_energy

### 2012-12-27 ParetoHMM/sim1.py

@@ -50,9 +50,19 @@ if __name__ == '__main__' :
 
 	pl.figure()
 	pl.plot(ll_list1,ll_list2,'b*')
-	pl.plot(*zip(*frontier_energy),color='magenta',marker='*',\
+	pl.plot(*zip(*sorted(frontier_energy)),color='magenta',marker='*',\
 		linestyle='dashed')
-	
+	ctr = dict(zip(set(frontier_energy),[0]*len(set(frontier_energy))))
+	for i,e in enumerate(frontier_energy) : 
+		ctr[e] += 1
+		if i == 0 :
+			pl.text(e[0]+0.5,e[1]-0.4,str(i),fontsize=10)
+			pl.text(e[0]-3,e[1]+0.8,frontier[i],fontsize=9)	
+		else :
+			pl.text(e[0]+0.5,e[1]-0.4,str(i),fontsize=10)
+			pl.text(e[0]-3,e[1]-1.5,frontier[i],fontsize=9)	
+
+
 	pl.xlabel('Energy:'+feat1)
 	pl.ylabel('Energy:'+feat2)
 	pl.title('Energy Plot')
@@ -63,3 +73,4 @@ if __name__ == '__main__' :
 	pl.axvline()
 	pl.axhline()
 	pl.savefig('../docs/tex/pics/sim1.pdf')
+	pl.show()

### 2012-12-27 ParetoHMM/sim3.py

@@ -1,7 +1,6 @@
 """
 Simulation Study 2 :
-Create a 12 length protein
-Run Pareto Frontier
+Run a number of different simulation studies
 
 @author: smoitra@cs.cmu.edu
 @license: BSD
@@ -36,6 +35,12 @@ def set_params_hmm_exp1(hmm) :
 	hmm.featmap = [{'H':0,'B':1,'L':2}]*hmm.length
 	hmm.initprob = [0.5,0.5]
 	hmm.trained = True
+
+class SimExp(object) :
+	""" Helps run a number of different simulation experiments """
+	def __init__(self)	:
+		pass
+
 
 if __name__ == '__main__' : 
 	hmm = HMM()
@@ -87,15 +92,15 @@ if __name__ == '__main__' :
 	# Plot all the points 
 	pl.figure()
 	pl.plot(ll_list1,ll_list2,'b*')
-	pl.plot(*zip(*sorted(frontier_energy)),color='magenta',marker='*',\
+	pl.plot(*sorted(zip(*frontier_energy)),color='magenta',marker='*',\
 		linestyle='dashed')
 	pl.xlabel('Energy:'+feat1)
 	pl.ylabel('Energy:'+feat2)
 	pl.title('Energy Plot')
 	xmin,xmax = pl.xlim()
 	ymin,ymax = pl.ylim()
-	pl.xlim(xmin,xmax)
-	pl.ylim(ymin,ymax)
+	pl.xlim(-2,xmax)
+	pl.ylim(-2,ymax)
 	pl.axvline()
 	pl.axhline()
 	pl.savefig('../docs/tex/pics/sim2_all.pdf')

### 2012-12-29 docs/tex/draft.tex
Additions to sim3 and changes to draft

@@
 
+Current benchmarking algorithms  : 
+\begin{itemize}
+\item Brute Force - Simply enumerate and find the frontier
+\item MCMC - Sample a number of sequences according to your energy function and keep the minimum. Should one apply Metropolis Hastings criteria usign Boltzmannm weights or something else
+\item Goldstein DEE - Need a way to assign pairwise energies between amino acids. 
+\item Yanover multistate DEE - Same as before but again need to assign weights. 
+\end{itemize}
+
+Also, multistate sequences designed by me are going to be good only so far as my emission and transmission weights are accurate.
+There needs to be a goldstandard way of assigning proper energies to the rotamers.
+
 \end{document}

### 2012-12-29 ParetoHMM/sim3.py

@@ -15,47 +15,51 @@ from cvxhull import pareto_frontier
 import random
 DEBUG=0
 
-def gen_random_dist(size) : 
-	x = np.random.uniform(size=size)
-	x = x/sum(x)
-	return x
-
-def set_params_hmm_exp1(hmm) : 
-	""" Sets the params of a hmm for sim experiment 1"""
-	hmm.length = 12
-	hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
-	hmm.emit = [
-		[gen_random_dist(3),gen_random_dist(3)]
-	]*hmm.length
-	hmm.trans = [
-		[gen_random_dist(2),gen_random_dist(2)]
-	]*hmm.length
-	hmm.seqmap = [{'a':0,'b':1}]*hmm.length
-	hmm.seqmap2 = [{0:'a',1:'b'}]*hmm.length
-	hmm.featmap = [{'H':0,'B':1,'L':2}]*hmm.length
-	hmm.initprob = [0.5,0.5]
-	hmm.trained = True
 
 class SimExp(object) :
 	""" Helps run a number of different simulation experiments """
-	def __init__(self)	:
-		pass
+	def __init__(self,name,ntimes,nlength=12,nstates=2,tied=True)	:
+		self.name = name #name of experiment
+		self.ntimes = ntimes # number of times to repeat the exp
+		hmm = HMM()		
+		set_params_hmm(hmm)
+		cmrf = CMRF(hmm)
 
+	def set_feats_standard(self) : 
+		self.feats = [
+			feat1 = 'HHHHLLLLHHHH'
+			feat2 = 'BBBBLLLLBBBB'
+		]
+
+
+	def gen_random_dist(size) : 
+		x = np.random.uniform(size=size)
+		x = x/sum(x)
+		return x
+	
+	def set_params_hmm(hmm) : 
+		""" Sets the params of a hmm for sim experiment 1"""
+		hmm.length = 12
+		hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
+		hmm.emit = [
+			[gen_random_dist(3),gen_random_dist(3)]
+		]*hmm.length
+		hmm.trans = [
+			[gen_random_dist(2),gen_random_dist(2)]
+		]*hmm.length
+		hmm.seqmap = [{'a':0,'b':1}]*hmm.length
[…]

### 2012-12-29 docs/tex/draft.tex
Designed and enumerated the simulation experiments that need to be rubn

@@
 
-\section{Experiments}
-In this section we evaluate the performance of our algorithm on simulated as well as experimental data.
-
-\subsection{Simulation Experiments}
-We run a variety of different simulation experiments to check the performance of our algorithm on
-
-\subsubsection{Sim Experiment 1}
-We demonstrate how our algorithm performs on simulated data.
-The data was prepared by considering a protein of length 12.
-The sequence space was limited to an alphabet consisting of two arbitrary characters $\alpha,\beta$.
-The observation belong to a discrete space Helix, Beta and Loop which represented by H,B and L.
-The weights of the CMRF have weights $\phi,\psi$ as defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2}.
-We define two states State A and State B that corresponding to competing objectives for multistate design.
-State A is a Helix-Loop-Helix strand whereas State B is Sheet-Loop-Sheet strand.
-Figure~\ref{fig:simstates} shows a cartoon diagrams for our design problem.
-State A and State B have \texttt{HHHHLLLLHHHH} and \texttt{BBBBLLLLBBBB} secondary structure features respectively.
-These are the observations for the CMRF.
-\\
-\\
-Using the weight matrices we defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2} we run CMRF-Decode to get $X_A$ which satisfies $X_A = \argmin_X E_A(X)$.
-We obtain $X_B$ similarly.
-Not surprisingly $X_A$ is $\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha$
-and $X_B$ is $\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta$.
-This is because the weights in the table are biased towards $\alpha$ emitting \texttt{H} and $\beta$ emitting \texttt{B}.
-Thus according to our model $X_A$ is the sequence that is most stable under state A, whereas $X_B$ is the sequence is most stable under state B.
-\\
-\\
-We then run the \texttt{Pareto-Frontier} algorithm using the same weights and observations for state A and state B to enumerate the entire pareto frontier.
-This pareto frontier contains all the best sequences that any convex combination of $E_A(X)$ and $E_B(X)$ would give rise to.
-We number the points to show the path taken by the pareto frontier algorithm.
-We also calculate the energies of all the $2^{12}$ sequences possible under this restricted sequence model and plot the points.
-The \texttt{Pareto-Frontier} algorithm was able to find the pareto frontier in \textcolor{red}{1 step} steps.
-See Figure~\ref{fig:pareto-sim} for the trace taken by the algorithm.
-We also compare the running times as opposed to brute force and MCMC and compare the running times.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cc|c|c|c|}
-\cline{3-5}
-& & \multicolumn{3}{c|}{Observations($\mathbf{O_i}$)} \\ \cline{3-5}
-& & Helix(H) & Beta(B) & Loop(L) \\ \cline{1-5}
-\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
-\multicolumn{1}{|c|}{$\alpha$} & 0.6 & 0.2 & 0.2      \\ \cline{2-5}
-\multicolumn{1}{|c|}{}                        &
-\multicolumn{1}{|c|}{$\beta$} & 0.2 & 0.6 & 0.2      \\ \cline{1-5}
-\end{tabular}
-\end{center}
-\caption{$\psi(X,\mathbf{O})$ contains the weights for edge potentials of sequence and features together }
-\label{tab:psi1}
-\end{table}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cc|c|c|}
-\cline{3-4}
-& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
-& & $\alpha$ & $\beta$  \\ \cline{1-4}
-\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
-\multicolumn{1}{|c|}{$\alpha$} & 0.7 & 0.3      \\ \cline{2-4}
-\multicolumn{1}{|c|}{}                        &
-\multicolumn{1}{|c|}{$\beta$} & 0.3 & 0.7      \\ \cline{1-4}
-\end{tabular}
-\end{center}
-\caption{$\psi(X_i,X_{i+1})$ contains the weights for the edge potentials of adjacent amino acids}
-\label{tab:psi2}
-\end{table}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cc|c|c|}
-\cline{3-4}
-& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
-& & $\alpha$ & $\beta$  \\ \cline{1-4}
-\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
-\multicolumn{1}{|c|}{$\alpha$} & 0.7 & 0.3      \\ \cline{2-4}
-\multicolumn{1}{|c|}{}                        &
-\multicolumn{1}{|c|}{$\beta$} & 0.3 & 0.7      \\ \cline{1-4}
-\end{tabular}
-\end{center}
-\caption{Time and steps taken on simulated data}
-\end{table}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{pics/sim1.pdf}
-    \caption{Simulation Experiment 1: The emission and transition weights are as defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2}.
-The Pareto-HMM finds the optimal pareto frontier for a twelve length sequence using the alphabet \texttt\{a,b\} and the feature space \texttt\{H,L,B\} in two steps of the Convex hull algorithm.
-The magenta stars corresponds to the optimal sequences.
-The magenta line corresponds to the lower left convex hull of the sequence energy space.
-The blue stars correspond to non-optimal dominated sequences.
-The Pareto-Frontier algoerithm provably finds the frontier quickly instead of having to enumerate,calculate and sort the energies of a  $2^{12}$ sequence space.}
-    \label{fig:pareto-sim}
-\end{figure}
-
-\pagebreak
-
-\subsubsection{Simulation Experiment 2}
-In this example we look at a slightly more complicated energy frontier.
-To generate this energy surface we resorted to generating random emission and transition probabilities till we found a sufficiently complicated energy frontier but still having alphabet \texttt{\{a,b\}} and feature space \texttt{\{H,L,B\}}.
-The probabilities for this simulation experiment are shown in tables~\ref{tab:psi3} and~\ref{tab:psi4}.
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cc|c|c|c|}
-\cline{3-5}
-& & \multicolumn{3}{c|}{Observations($\mathbf{O_i}$)} \\ \cline{3-5}
-& & Helix(H) & Beta(B) & Loop(L) \\ \cline{1-5}
-\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
-\multicolumn{1}{|c|}{$\alpha$} & 0.334 & 0.272 & 0.394      \\ \cline{2-5}
-\multicolumn{1}{|c|}{}                        &
-\multicolumn{1}{|c|}{$\beta$} & 0.477 & 0.093 & 0.430      \\ \cline{1-5}
-\end{tabular}
-\end{center}
-\caption{$\psi(X,\mathbf{O})$ contains the weights for edge potentials of sequence and features together }
-\label{tab:psi3}
-\end{table}
-
-\begin{table}
-\begin{center}
-\begin{tabular}{cc|c|c|}
-\cline{3-4}
-& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
-& & $\alpha$ & $\beta$  \\ \cline{1-4}
-\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
-\multicolumn{1}{|c|}{$\alpha$} & 0.483 & 0.517      \\ \cline{2-4}
-\multicolumn{1}{|c|}{}                        &
-\multicolumn{1}{|c|}{$\beta$} & 0.589 & 0.411      \\ \cline{1-4}
-\end{tabular}
-\end{center}
-\caption{$\psi(X_i,X_{i+1})$ contains the weights for the edge potentials of adjacent amino acids}
-\label{tab:psi4}
-\end{table}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{pics/sim2_all.pdf}
-    \caption{Simulation Experiment 2: Pareto Frontier algorithm is applied to a more complicated energy surface.
-The algorithm finds the pareto frontier denoted by the dashed magenta in 5 steps, which is significantly faster than brute force enumeration.
-}
-    \label{fig:sim2_all}
-\end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[width=0.9\textwidth]{pics/sim2_frontier.pdf}
-    \caption{Simulation Experiment 2(Frontier): This shows the results from simulation 2. The sequences along the frontier are shown as well }
-    \label{fig:sim2_frontier}
-\end{figure}
-
-\pagebreak
-\clearpage
-
-\subsection{Experiments on Biological Data}
+\include{experiments}
 

### 2012-12-29 docs/tex/experiments.tex
Designed and enumerated the simulation experiments that need to be rubn

\section{Experiments}
In this section we evaluate the performance of our algorithm on simulated as well as experimental data.

\subsection{Simulation Experiments}
We run a variety of different simulation experiments to check the performance of our algorithm on

\subsubsection{Sim Experiment 1}
\label{sim:toy}
We demonstrate how our algorithm performs on simulated data. The data was prepared by considering a protein of length 12. The sequence space was limited to an alphabet consisting of two arbitrary characters $\alpha,\beta$. The observation belong to a discrete space Helix, Beta and Loop which represented by H,B and L. The weights of the CMRF have weights $\phi,\psi$ as defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2}. We define two states State A and State B that corresponding to competing objectives for multistate design. State A is a Helix-Loop-Helix strand whereas State B is Sheet-Loop-Sheet strand. Figure~\ref{fig:simstates} shows a cartoon diagrams for our design problem. State A and State B have \texttt{HHHHLLLLHHHH} and \texttt{BBBBLLLLBBBB} secondary structure features respectively. These are the observations for the CMRF.
\\
\\
Using the weight matrices we defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2} we run CMRF-Decode to get $X_A$ which satisfies $X_A = \argmin_X E_A(X)$. We obtain $X_B$ similarly. Not surprisingly $X_A$ is $\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha\alpha$ 
and $X_B$ is $\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta\beta$. This is because the weights in the table are biased towards $\alpha$ emitting \texttt{H} and $\beta$ emitting \texttt{B}. Thus according to our model $X_A$ is the sequence that is most stable under state A, whereas $X_B$ is the sequence is most stable under state B.
\\
\\
We then run the \texttt{Pareto-Frontier} algorithm using the same weights and observations for state A and state B to enumerate the entire pareto frontier. This pareto frontier contains all the best sequences that any convex combination of $E_A(X)$ and $E_B(X)$ would give rise to. We number the points to show the path taken by the pareto frontier algorithm. We also calculate the energies of all the $2^{12}$ sequences possible under this restricted sequence model and plot the points. The \texttt{Pareto-Frontier} algorithm was able to find the pareto frontier in \textcolor{red}{1 step} steps. See Figure~\ref{fig:pareto-sim} for the trace taken by the algorithm.  We also compare the running times as opposed to brute force and MCMC and compare the running times.

\begin{table}
\begin{center}
\begin{tabular}{cc|c|c|c|}
\cline{3-5}
& & \multicolumn{3}{c|}{Observations($\mathbf{O_i}$)} \\ \cline{3-5}
& & Helix(H) & Beta(B) & Loop(L) \\ \cline{1-5}
\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
\multicolumn{1}{|c|}{$\alpha$} & 0.6 & 0.2 & 0.2      \\ \cline{2-5}
\multicolumn{1}{|c|}{}                        &
\multicolumn{1}{|c|}{$\beta$} & 0.2 & 0.6 & 0.2      \\ \cline{1-5}
\end{tabular}
\end{center}
\caption{$\psi(X,\mathbf{O})$ contains the weights for edge potentials of sequence and features together }
\label{tab:psi1}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{cc|c|c|}
\cline{3-4}
& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
& & $\alpha$ & $\beta$  \\ \cline{1-4}
\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
\multicolumn{1}{|c|}{$\alpha$} & 0.7 & 0.3      \\ \cline{2-4}
\multicolumn{1}{|c|}{}                        &
\multicolumn{1}{|c|}{$\beta$} & 0.3 & 0.7      \\ \cline{1-4}
\end{tabular}
\end{center}
\caption{$\psi(X_i,X_{i+1})$ contains the weights for the edge potentials of adjacent amino acids}
\label{tab:psi2}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{cc|c|c|}
\cline{3-4}
& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
& & $\alpha$ & $\beta$  \\ \cline{1-4}
\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
\multicolumn{1}{|c|}{$\alpha$} & 0.7 & 0.3      \\ \cline{2-4}
\multicolumn{1}{|c|}{}                        &
\multicolumn{1}{|c|}{$\beta$} & 0.3 & 0.7      \\ \cline{1-4}
\end{tabular}
\end{center}
\caption{Time and steps taken on simulated data}
\end{table}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{pics/sim1.pdf}
    \caption{Simulation Experiment 1: The emission and transition weights are as defined in Tables~\ref{tab:psi1} and~\ref{tab:psi2}. The Pareto-HMM finds the optimal pareto frontier for a twelve length sequence using the alphabet \texttt\{a,b\} and the feature space \texttt\{H,L,B\} in two steps of the Convex hull algorithm. The magenta stars corresponds to the optimal sequences. The magenta line corresponds to the lower left convex hull of the sequence energy space. The blue stars correspond to non-optimal dominated sequences.  The Pareto-Frontier algoerithm provably finds the frontier quickly instead of having to enumerate,calculate and sort the energies of a  $2^{12}$ sequence space.}
    \label{fig:pareto-sim}
\end{figure}

\pagebreak

\subsubsection{Simulation Experiment 2}
\label{sim:randprobs}
In this example we look at a slightly more complicated energy frontier. To generate this energy surface we resorted to generating random emission and transition probabilities till we found a sufficiently complicated energy frontier but still having alphabet \texttt{\{a,b\}} and feature space \texttt{\{H,L,B\}}. The probabilities for this simulation experiment are shown in tables~\ref{tab:psi3} and~\ref{tab:psi4}.

\begin{table}
\begin{center}
\begin{tabular}{cc|c|c|c|}
\cline{3-5}
& & \multicolumn{3}{c|}{Observations($\mathbf{O_i}$)} \\ \cline{3-5}
& & Helix(H) & Beta(B) & Loop(L) \\ \cline{1-5}
\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
\multicolumn{1}{|c|}{$\alpha$} & 0.334 & 0.272 & 0.394      \\ \cline{2-5}
\multicolumn{1}{|c|}{}                        &
\multicolumn{1}{|c|}{$\beta$} & 0.477 & 0.093 & 0.430      \\ \cline{1-5}
\end{tabular}
\end{center}
\caption{$\psi(X,\mathbf{O})$ contains the weights for edge potentials of sequence and features together }
\label{tab:psi3}
\end{table}

\begin{table}
\begin{center}
\begin{tabular}{cc|c|c|}
\cline{3-4}
& & \multicolumn{2}{c|}{Sequence($X_{i+1}$)} \\ \cline{3-4}
& & $\alpha$ & $\beta$  \\ \cline{1-4}
\multicolumn{1}{|c|}{\multirow{2}{*}{Sequence($X_i$)}} &
\multicolumn{1}{|c|}{$\alpha$} & 0.483 & 0.517      \\ \cline{2-4}
\multicolumn{1}{|c|}{}                        &
\multicolumn{1}{|c|}{$\beta$} & 0.589 & 0.411      \\ \cline{1-4}
\end{tabular}
\end{center}
\caption{$\psi(X_i,X_{i+1})$ contains the weights for the edge potentials of adjacent amino acids}
\label{tab:psi4}
\end{table}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{pics/sim2_all.pdf}
    \caption{Simulation Experiment 2: Pareto Frontier algorithm is applied to a more complicated energy surface. The algorithm finds the pareto frontier denoted by the dashed magenta in 5 steps, which is significantly faster than brute force enumeration.  }
    \label{fig:sim2_all}
\end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.9\textwidth]{pics/sim2_frontier.pdf}
    \caption{Simulation Experiment 2(Frontier): This shows the results from simulation 2. The sequences along the frontier are shown as well }
    \label{fig:sim2_frontier}
\end{figure}

\pagebreak
\clearpage

\subsubsection{Simulation Experiment 3: Random probs tied}
\label{sim:randprobstied}
In this experiment we use a 12 length sequence. We do 2-state design using features \texttt{HHHHLLLLHHHH} and \texttt{BBBBLLLLBBBB} as before. The sequence space is restricted to a dummy \texttt{\{a,b\}}  We randomize the emission and transition probabilities. The parameters of the emission and the transition are ties together.  We store the frontier, frontier energies, computation time taken and

\subsubsection{Simulation Experiment 4:Random probs untied}
\label{sim:randprobsuntied}
Same as Simulation~\ref{sim:randprobstied} with the difference that the emission and the transition probabilities are untied.

\subsubsection{Simulation Experiment 5 : Random feature space tied}

\subsubsection{Simulation Experiment 6 :  Random feature space untied}

\subsubsection{Simulation Experiment 7: Vary Seq space }

\subsubsection{Simulation Experiment 8: Seq space 20, Vary seq length}

\subsubsection{Simulation Experiment 9 : Jointly vary seq space and length}

\subsubsection{Simulation Experiment 10 : K-way multistate design}

\subsection{Benchmarking with other Methods}

\subsection{Experiments on Biological Data}

### 2012-12-29 ParetoHMM/sim3.py

@@ -18,7 +18,8 @@ DEBUG=0
 
 class SimExp(object) :
 	""" Helps run a number of different simulation experiments """
-	def __init__(self,name,ntimes,nlength=12,nstates=2,tied=True)	:
+	def __init__(self,name,ntimes=10,nlength=12,nstates=2,tied=True\
+		)	:
 		self.name = name #name of experiment
 		self.ntimes = ntimes # number of times to repeat the exp
 		hmm = HMM()		
@@ -30,8 +31,6 @@ class SimExp(object) :
 			feat1 = 'HHHHLLLLHHHH'
 			feat2 = 'BBBBLLLLBBBB'
 		]
-
-
 	def gen_random_dist(size) : 
 		x = np.random.uniform(size=size)
 		x = x/sum(x)

### 2012-12-30 ParetoHMM/sim3.py
refactored some sim3 code

@@ -18,13 +18,34 @@ DEBUG=0
 
 class SimExp(object) :
 	""" Helps run a number of different simulation experiments """
-	def __init__(self,name,ntimes=10,nlength=12,nstates=2,tied=True\
+	def __init__(self,name,ntimes=10,length=12,seqstates=2,tied=True\
 		)	:
+		""" Set up the simulation study according to the parameters 
+		specified"""
 		self.name = name #name of experiment
 		self.ntimes = ntimes # number of times to repeat the exp
-		hmm = HMM()		
-		set_params_hmm(hmm)
-		cmrf = CMRF(hmm)
+		self.length = length
+		self.seqstates = seqstates
+		self.tied = tied
+		self.hmm = HMM()	
+		self.namemap = {
+		'toy' : self.toy
+		'randprobs' : self.randprobs
+		'randprobstied' : self.randprobstied
+		}	
+		self.namemap[self.name]()
+		self.execute()
+
+	def toy(self) :
+		""" Set up the toy simulation """
+		pass
+
+	def randprobs(self) : 
+		""" Set up a version of toy with non-trivial surface """
+		pass
+
+	def randprobstied(self) : 
+		""" Run many iterations of toy with random probs  """
 
 	def set_feats_standard(self) : 
 		self.feats = [
@@ -52,13 +73,27 @@ class SimExp(object) :
 		hmm.initprob = [0.5,0.5]
 		hmm.trained = True
 
-if __name__ == '__main__' : 
-	hmm = HMM()
-	# Set the params of the h,,
+	def bruteforce(self)  :
+		""" Run Brute force enumeration of the sequence space """
+		feat1,feat2 = self.feats
+		ll_list1,ll_list2 = [],[]
+		
+		seq_list = ["".join(s) for s in product('ab',repeat=12)]
+		for seq in seq_list:	
+			ll_list1.append(cmrf.score(seq,feat1))
+			ll_list2.append(cmrf.score(seq,feat2))
+	
+		min_feat1id = ll_list1.index(min(ll_list1))
+		min_feat2id = ll_list2.index(min(ll_list2))
 
-	### DEBUG
-	import pickle
-	cmrf = pickle.load(open('cmrf.pkl'))
+
+	def mcmc(self) : 
+		""" Run mcmc on the same problem instance """
+		### TODO
+		pass
+
+
+if __name__ == '__main__' : 
 
 	# Plot the entire sequence space
 	ll_list1,ll_list2 = [],[]

### 2013-01-03 ParetoHMM/sim3.py
Created simulation framework for toy experiments

@@ -13,13 +13,17 @@ from HMM import HMM,CMRF,TMRF
 from itertools import product
 from cvxhull import pareto_frontier 
 import random
+import time
+import pickle
 DEBUG=0
+
+STUDY='3'
 
 
 class SimExp(object) :
 	""" Helps run a number of different simulation experiments """
 	def __init__(self,name,ntimes=10,length=12,seqstates=2,tied=True\
-		)	:
+		,plot_all=False)	:
 		""" Set up the simulation study according to the parameters 
 		specified"""
 		self.name = name #name of experiment
@@ -27,18 +31,67 @@ class SimExp(object) :
 		self.length = length
 		self.seqstates = seqstates
 		self.tied = tied
-		self.hmm = HMM()	
-		self.namemap = {
-		'toy' : self.toy
-		'randprobs' : self.randprobs
+		self.plot_all = plot_all
+
+		# Set the random seed
+	
+	def execute(self)  :
+		""" Execute the stored execution function """	
+		# Set the random seed so that all experiments are pseudo random
+		random.seed(42)
+		namemap = {
+		'toy' : self.toy,
+		'randprobs' : self.randprobs,
 		'randprobstied' : self.randprobstied
 		}	
-		self.namemap[self.name]()
-		self.execute()
+		namemap[self.name]()
+		with open('data/sim'+str(STUDY)+'_'+self.name+'.pkl','w') as \
+			pklout : 
+			pickle.dump(self,pklout)
 
 	def toy(self) :
-		""" Set up the toy simulation """
-		pass
+		""" Set up the toy simulation """	
+		self.tasklist = []
+		feats = self.get_feats_standard()
+		hmm = HMM()
+		self._set_params_toy(hmm)
+		cmrf = CMRF(hmm)
+		for taskid in range(self.ntimes) :	
+			task = Task('sim'+STUDY+'_'+self.name+'_'+str(taskid),cmrf,\
+				feats)				
+			# Run Brute force to enumerate the frontier
+			with benchmark(task.name+'brute') as t:
+				seq,energies = self.bruteforce(cmrf,feats)			
+			task.all_seq = seq
[…]

### 2013-01-03 ParetoHMM/sim3.py
Added sim 2

@@ -95,7 +95,64 @@ class SimExp(object) :
 
 	def randprobs(self) : 
 		""" Set up a version of toy with non-trivial surface """
-		pass
+		self.tasklist = []
+		feats = self.get_feats_standard()
+		hmm = HMM()
+		self._set_params_randprobs(hmm)
+		cmrf = CMRF(hmm)
+		for taskid in range(self.ntimes) :	
+			task = Task('sim'+STUDY+'_'+self.name+'_'+str(taskid),cmrf,\
+				feats)				
+			# Run Brute force to enumerate the frontier
+			with benchmark(task.name+'brute') as t:
+				seq,energies = self.bruteforce(cmrf,feats)			
+			task.all_seq = seq
+			task.all_seq_energy = energies
+			task.brute_time = t.elapsed			
+
+			# Now run the toy simulation`
+			with benchmark(task.name+'pareto') as t : 
+				task.frontier,task.frontier_energy = \
+					pareto_frontier(cmrf,feats)		
+			if self.plot_all :
+				task.plot_frontier()
+			task.pareto_time = t.elapsed
+			self.tasklist.append(task)	
+
+	def _set_params_randprobs(self,hmm) :
+		""" Sets the params of a hmm for sim experiment 1"""
+		hmm.length = 12
+		hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
+		hmm.emit = [
+			[[0.334,0.272,0.394],[0.477,0.093,0.430]]
+		]*hmm.length
+		hmm.trans = [
+			[[0.483,0.517],[0.589,0.411]]
+		]*hmm.length
+
+		hmm.seqmap = [{'a':0,'b':1}]*hmm.length
+		hmm.seqmap2 = [{0:'a',1:'b'}]*hmm.length
+		hmm.featmap = [{'H':0,'B':1,'L':2}]*hmm.length
+		hmm.initprob = [0.5,0.5]
+		hmm.trained = True
+
+
+	def set_params_randprobstied(hm) : 
+		""" Sets the params of a hmm for sim experiment 1"""
+		hmm.length = 12
+		hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
+		hmm.emit = [
+			[gen_random_dist(3),gen_random_dist(3)]
+		]*hmm.length
+		hmm.trans = [
[…]

### 2013-01-03 ParetoHMM/sim3.py
Set up randprobs tied

@@ -39,6 +39,7 @@ class SimExp(object) :
 		""" Execute the stored execution function """	
 		# Set the random seed so that all experiments are pseudo random
 		random.seed(42)
+		np.random.seed(42)
 		namemap = {
 		'toy' : self.toy,
 		'randprobs' : self.randprobs,
@@ -137,26 +138,49 @@ class SimExp(object) :
 		hmm.trained = True
 
 
-	def set_params_randprobstied(hm) : 
+	def randprobstied(self) : 
+		""" Run many iterations of toy with random probs  """
+		self.tasklist = []
+		feats = self.get_feats_standard()
+	
+		# Repeat for all the tasks described	
+		for taskid in range(self.ntimes) :	
+			hmm = HMM()
+			self._set_params_randprobstied(hmm)
+			cmrf = CMRF(hmm)	
+			task = Task('sim'+STUDY+'_'+self.name+'_'+str(taskid),cmrf,\
+				feats)				
+			# Run Brute force to enumerate the frontier
+			with benchmark(task.name+'brute') as t:
+				seq,energies = self.bruteforce(cmrf,feats)			
+			task.all_seq = seq
+			task.all_seq_energy = energies
+			task.brute_time = t.elapsed			
+
+			# Now run the toy simulation`
+			with benchmark(task.name+'pareto') as t : 
+				task.frontier,task.frontier_energy = \
+					pareto_frontier(cmrf,feats)		
+			if self.plot_all :
+				task.plot_frontier()
+			task.pareto_time = t.elapsed
+			self.tasklist.append(task)	
+
+	def _set_params_randprobstied(self,hmm) : 
 		""" Sets the params of a hmm for sim experiment 1"""
 		hmm.length = 12
 		hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
 		hmm.emit = [
-			[gen_random_dist(3),gen_random_dist(3)]
+			[self.gen_random_dist(3),self.gen_random_dist(3)]
 		]*hmm.length
 		hmm.trans = [
-			[gen_random_dist(2),gen_random_dist(2)]
+			[self.gen_random_dist(2),self.gen_random_dist(2)]
 		]*hmm.length
 		hmm.seqmap = [{'a':0,'b':1}]*hmm.length
 		hmm.seqmap2 = [{0:'a',1:'b'}]*hmm.length
[…]

### 2013-01-03 docs/tex/experiments.tex
Added random features tied and untied

@@
 \subsubsection{Simulation Experiment 5 : Random feature space tied}
+Like~\ref{sim:randprobstied} but the feature space is random
+\label{sim:randfeatstied}
 
 \subsubsection{Simulation Experiment 6 :  Random feature space untied}
+Like~\ref{sim:randprobsuntied} but the feature space is random
+\label{sim:randfeatsuntied}
 

### 2013-01-03 ParetoHMM/sim3.py

@@ -43,7 +43,10 @@ class SimExp(object) :
 		namemap = {
 		'toy' : self.toy,
 		'randprobs' : self.randprobs,
-		'randprobstied' : self.randprobstied
+		'randprobstied' : self.randprobstied,
+		'randprobsuntied' : self.randprobsuntied,
+		'randfeatstied' : self.randfeatstied,
+		'randfeatsuntied' : self.randfeatsuntied
 		}	
 		namemap[self.name]()
 		with open('data/sim'+str(STUDY)+'_'+self.name+'.pkl','w') as \
@@ -182,6 +185,118 @@ class SimExp(object) :
 		hmm.initprob = [0.5,0.5]
 		hmm.trained = True
 
+	def randprobsuntied(self) : 
+		""" Run many iterations of toy with random probs  """
+		self.tasklist = []
+		feats = self.get_feats_standard()
+	
+		# Repeat for all the tasks described	
+		for taskid in range(self.ntimes) :	
+			hmm = HMM()
+			self._set_params_randprobsuntied(hmm)
+			cmrf = CMRF(hmm)	
+			task = Task('sim'+STUDY+'_'+self.name+'_'+str(taskid),cmrf,\
+				feats)				
+			# Run Brute force to enumerate the frontier
+			with benchmark(task.name+'brute') as t:
+				seq,energies = self.bruteforce(cmrf,feats)			
+			task.all_seq = seq
+			task.all_seq_energy = energies
+			task.brute_time = t.elapsed			
+
+			# Now run the toy simulation`
+			with benchmark(task.name+'pareto') as t : 
+				task.frontier,task.frontier_energy = \
+					pareto_frontier(cmrf,feats)		
+			if self.plot_all :
+				task.plot_frontier()
+			task.pareto_time = t.elapsed
+			self.tasklist.append(task)	
+
+	def _set_params_randprobsuntied(self,hmm) : 
+		""" Sets the params of a hmm for sim experiment 1"""
+		hmm.length = 12
+		hmm.dims = [(2,3)]*hmm.length # (latent,emit) dimspace
+		hmm.emit,hmm.trans = [],[]
+		for i in range(hmm.length) : 
+			hmm.emit.append(
+				[self.gen_random_dist(3),self.gen_random_dist(3)]
+			)
+			hmm.trans.append(
+				[self.gen_random_dist(2),self.gen_random_dist(2)]
+			)
+		hmm.seqmap = [{'a':0,'b':1}]*hmm.length
[…]

### 2013-01-03 docs/tex/experiments.tex
Simulation 7, seqspace

@@
 \subsubsection{Simulation Experiment 7: Vary Seq space }
+\label{sim:seqspace}
+Increase the sequence space using random untied feature space as in ~\ref{sim:randfeatsuntied}. Will supply different seq space values. 2,4,16,20 are the values that are goint to be used.
 
 \subsubsection{Simulation Experiment 8: Seq space 20, Vary seq length}
+\label{sim:seqlen}
+Set the sequence space to be 20 and vary the sequence length from 2^2,2^16
 
 \subsubsection{Simulation Experiment 9 : Jointly vary seq space and length}
+\label{sim:seqspacelen}
+Jointly vary the sequence space and sequence length,
 
-\subsubsection{Simulation Experiment 10 : K-way multistate design}
+\subsubsection{Simulation Experiment 10: Vary Feat space }
+\label{sim:featspace}
+Keeping the seq space at 20, vary feat space from 5-10-25-50
+
+\subsubsection{Simulation Experiment 11: Vary Feat space and sequence length}
+\label{sim:featspacelen}
+Keeping the seq space at 20, vary feat space from 5-10-25-50, sequence length from 2^2,2^16
+
+\subsubsection{Simulation Experiment : K-way multistate design}
+
+\subsection{Simulation Experiment 11 : Augmenting Feature space}
 

### 2013-01-03 ParetoHMM/sim3.py

@@ -23,7 +23,7 @@ STUDY='3'
 class SimExp(object) :
 	""" Helps run a number of different simulation experiments """
 	def __init__(self,name,ntimes=10,length=12,seqstates=2,tied=True\
-		,plot_all=False)	:
+		,plot_all=False,**kwdargs)	:
 		""" Set up the simulation study according to the parameters 
 		specified"""
 		self.name = name #name of experiment
@@ -32,7 +32,7 @@ class SimExp(object) :
 		self.seqstates = seqstates
 		self.tied = tied
 		self.plot_all = plot_all
-
+		self.kwdargs = kwdargs
 		# Set the random seed
 	
 	def execute(self)  :
@@ -46,7 +46,12 @@ class SimExp(object) :
 		'randprobstied' : self.randprobstied,
 		'randprobsuntied' : self.randprobsuntied,
 		'randfeatstied' : self.randfeatstied,
-		'randfeatsuntied' : self.randfeatsuntied
+		'randfeatsuntied' : self.randfeatsuntied,
+		'seqspace' : self.seqspace , 
+		'seqlen' :  self.seqlen , 
+		'seqspacelen': self.seqspacelen,
+		'featspace' : self.featspace,
+		'featspacelen' : self.featspacelen
 		}	
 		namemap[self.name]()
 		with open('data/sim'+str(STUDY)+'_'+self.name+'.pkl','w') as \
@@ -289,12 +294,96 @@ class SimExp(object) :
 			task.pareto_time = t.elapsed
 			self.tasklist.append(task)	
 
+	def _set_params_generic(self,hmm,length,dims) : 
+		""" Sets the params of a hmm for sim experiment """
+		hmm.length = length
+		hmm.dims = dims # (latent,emit) dimspace
+		hmm.emit,hmm.trans = [],[]
+		for i in range(hmm.length) : 
+			hmm.emit.append(
+				[self.gen_random_dist(dims[i][1]) \
+					for j in range(dims[i][0])]
+			)
+			hmm.trans.append(
+				[self.gen_random_dist(dims[i][0]) \
+					for j in range(dims[i][0])]
+			)
+		seqspace = 'ACDEFGHIKLMNPQRSTVWY'.lower()
+		seqdict2 = dict(enumerate(seqspace))
+		seqdict = dict([(a,i) for i,a in enumerate(seqspace)])
[…]

### 2013-01-03 docs/tex/experiments.tex
Finished exp9. jointly varying seq space and length. Added timing information too

@@
 \label{sim:randfeatsuntied}
+sim3_toy_0brute : 0.907 seconds
+sim3_toy_0pareto : 0.003 seconds
+sim3_randprobs_0brute : 0.986 seconds
+sim3_randprobs_0pareto : 0.006 seconds
+sim3_randprobstied_0brute : 0.912 seconds
+sim3_randprobstied_0pareto : 0.004 seconds
+sim3_randprobsuntied_0brute : 0.903 seconds
+sim3_randprobsuntied_0pareto : 0.004 seconds
+sim3_randfeatstied_0brute : 0.921 seconds
+sim3_randfeatstied_0pareto : 0.011 seconds
+sim3_randfeatsuntied_0brute : 0.910 seconds
+sim3_randfeatsuntied_0pareto : 0.006 seconds
 
@@
 Increase the sequence space using random untied feature space as in ~\ref{sim:randfeatsuntied}. Will supply different seq space values. 2,4,16,20 are the values that are goint to be used.
+sim3_seqspace_2_0pareto : 0.017 seconds
+sim3_seqspace_2_1pareto : 0.005 seconds
+sim3_seqspace_2_2pareto : 0.007 seconds
+sim3_seqspace_4_0pareto : 0.022 seconds
+sim3_seqspace_4_1pareto : 0.021 seconds
+sim3_seqspace_4_2pareto : 0.015 seconds
+sim3_seqspace_8_0pareto : 0.040 seconds
+sim3_seqspace_8_1pareto : 0.069 seconds
+sim3_seqspace_8_2pareto : 0.061 seconds
+sim3_seqspace_16_0pareto : 0.338 seconds
+sim3_seqspace_16_1pareto : 0.213 seconds
+sim3_seqspace_16_2pareto : 0.214 seconds
+sim3_seqspace_20_0pareto : 0.370 seconds
+sim3_seqspace_20_1pareto : 0.463 seconds
+sim3_seqspace_20_2pareto : 0.284 seconds
 
@@
 Set the sequence space to be 20 and vary the sequence length from 2^2,2^16
+sim3_seqlen_4_0pareto : 0.047 seconds
+sim3_seqlen_4_1pareto : 0.011 seconds
+sim3_seqlen_4_2pareto : 0.029 seconds
+sim3_seqlen_16_0pareto : 0.621 seconds
+sim3_seqlen_16_1pareto : 0.857 seconds
+sim3_seqlen_16_2pareto : 0.499 seconds
+sim3_seqlen_32_0pareto : 2.343 seconds
+sim3_seqlen_32_1pareto : 2.344 seconds
+sim3_seqlen_32_2pareto : 2.409 seconds
+sim3_seqlen_64_0pareto : 8.683 seconds
+sim3_seqlen_64_1pareto : 10.716 seconds
+sim3_seqlen_64_2pareto : 11.068 seconds
+sim3_seqlen_128_0pareto : 39.451 seconds
+sim3_seqlen_128_1pareto : 40.050 seconds
+sim3_seqlen_128_2pareto : 33.411 seconds
 
@@
 Jointly vary the sequence space and sequence length, 
+sim3_seqspacelen_4_2_0pareto : 0.001 seconds
+sim3_seqspacelen_4_2_1pareto : 0.001 seconds
+sim3_seqspacelen_4_4_0pareto : 0.000 seconds
+sim3_seqspacelen_4_4_1pareto : 0.004 seconds
+sim3_seqspacelen_4_8_0pareto : 0.010 seconds
+sim3_seqspacelen_4_8_1pareto : 0.003 seconds
+sim3_seqspacelen_4_16_0pareto : 0.032 seconds
+sim3_seqspacelen_4_16_1pareto : 0.014 seconds
+sim3_seqspacelen_16_2_0pareto : 0.019 seconds
+sim3_seqspacelen_16_2_1pareto : 0.017 seconds
+sim3_seqspacelen_16_4_0pareto : 0.040 seconds
+sim3_seqspacelen_16_4_1pareto : 0.039 seconds
+sim3_seqspacelen_16_8_0pareto : 0.144 seconds
+sim3_seqspacelen_16_8_1pareto : 0.102 seconds
+sim3_seqspacelen_16_16_0pareto : 0.375 seconds
+sim3_seqspacelen_16_16_1pareto : 0.398 seconds
+sim3_seqspacelen_32_2_0pareto : 0.012 seconds
+sim3_seqspacelen_32_2_1pareto : 0.044 seconds
+sim3_seqspacelen_32_4_0pareto : 0.109 seconds
+sim3_seqspacelen_32_4_1pareto : 0.156 seconds
+sim3_seqspacelen_32_8_0pareto : 0.290 seconds
+sim3_seqspacelen_32_8_1pareto : 0.389 seconds
+sim3_seqspacelen_32_16_0pareto : 1.166 seconds
+sim3_seqspacelen_32_16_1pareto : 1.821 seconds
+sim3_seqspacelen_64_2_0pareto : 0.195 seconds
+sim3_seqspacelen_64_2_1pareto : 0.191 seconds
+sim3_seqspacelen_64_4_0pareto : 0.567 seconds
+sim3_seqspacelen_64_4_1pareto : 0.458 seconds
+sim3_seqspacelen_64_8_0pareto : 1.594 seconds
+sim3_seqspacelen_64_8_1pareto : 1.679 seconds
+sim3_seqspacelen_64_16_0pareto : 7.087 seconds
+sim3_seqspacelen_64_16_1pareto : 5.645 seconds
+sim3_seqspacelen_128_2_0pareto : 0.713 seconds
+sim3_seqspacelen_128_2_1pareto : 0.670 seconds
+sim3_seqspacelen_128_4_0pareto : 2.135 seconds
+sim3_seqspacelen_128_4_1pareto : 2.418 seconds
+sim3_seqspacelen_128_8_0pareto : 6.247 seconds
+sim3_seqspacelen_128_8_1pareto : 6.642 seconds
+sim3_seqspacelen_128_16_0pareto : 23.922 seconds
+sim3_seqspacelen_128_16_1pareto : 24.432 seconds
 

### 2013-01-03 ParetoHMM/sim3.py

@@ -54,7 +54,20 @@ class SimExp(object) :
 		'featspacelen' : self.featspacelen
 		}	
 		namemap[self.name]()
-		with open('data/sim'+str(STUDY)+'_'+self.name+'.pkl','w') as \
+		if self.name == 'seqspace' : 
+			fname = 'data/sim'+str(STUDY)+'_'+self.name+'_'+\
+				str(self.kwdargs['seqspace'])+'.pkl'
+		elif  self.name == 'seqlen' : 
+			fname = 'data/sim'+str(STUDY)+'_'+self.name+'_'+\
+				str(self.kwdargs['seqlen'])+'.pkl'
+		elif  self.name == 'seqspacelen' : 
+			fname = 'data/sim'+str(STUDY)+'_'+self.name+'_'+\
+				'_'+str(self.kwdargs['seqspace'])+\
+				'_'+str(self.kwdargs['seqlen'])+'.pkl'
+		else :
+			fname = 'data/sim'+str(STUDY)+'_'+self.name+'.pkl'
+
+		with open(fname,'w') as \
 			pklout : 
 			pickle.dump(self,pklout)
 
@@ -322,7 +335,6 @@ class SimExp(object) :
 	def seqspace(self) :
 		""" Vary sequence space """
 		self.tasklist = []
-		feats = self.get_feats_standard()
 		featspace = 3
 		seqlen = 12
 		dims = [(self.kwdargs['seqspace'],featspace)]*seqlen
@@ -337,28 +349,86 @@ class SimExp(object) :
 				str(self.kwdargs['seqspace'])+\
 				'_'+str(taskid),cmrf,feats)				
 			# Run Brute force to enumerate the frontier
-			with benchmark(task.name+'brute') as t:
-				seq,energies = self.bruteforce(cmrf,feats)			
-			task.all_seq = seq
-			task.all_seq_energy = energies
-			task.brute_time = t.elapsed			
+			if self.kwdargs['run_brute'] : 
+				with benchmark(task.name+'brute') as t:
+					seq,energies = self.bruteforce(cmrf,feats)			
+				task.all_seq = seq
+				task.all_seq_energy = energies
+				task.brute_time = t.elapsed			
 
 			# Now run the toy simulation`
 			with benchmark(task.name+'pareto') as t : 
 				task.frontier,task.frontier_energy = \
 					pareto_frontier(cmrf,feats)		
 			if self.plot_all :
-				task.plot_frontier()
[…]

### 2013-01-03 docs/tex/experiments.tex
Experiment 11. Feature space and sequence length

@@
 \label{sim:featspace}
-Keeping the seq space at 20, vary feat space from 5-10-25-50
+Keeping the seq space at 20 and seq len at 16. vary feat space from 5-10-25-50
+
+sim3_featspace_5_0pareto : 0.344 seconds
+sim3_featspace_5_1pareto : 0.426 seconds
+sim3_featspace_5_2pareto : 0.499 seconds
+sim3_featspace_10_0pareto : 0.543 seconds
+sim3_featspace_10_1pareto : 0.706 seconds
+sim3_featspace_10_2pareto : 0.657 seconds
+sim3_featspace_25_0pareto : 0.632 seconds
+sim3_featspace_25_1pareto : 0.493 seconds
+sim3_featspace_25_2pareto : 0.536 seconds
+sim3_featspace_50_0pareto : 0.543 seconds
+sim3_featspace_50_1pareto : 0.676 seconds
+sim3_featspace_50_2pareto : 0.586 seconds
 
@@
 \label{sim:featspacelen}
-Keeping the seq space at 20, vary feat space from 5-10-25-50, sequence length from 2^2,2^16
+Keeping the seq space at 20, vary feat space from 5-10-25-50, sequence length from seqlen to 64.
+sim3_featspacelen_4_5_0pareto : 0.030 seconds
+sim3_featspacelen_4_5_1pareto : 0.037 seconds
+sim3_featspacelen_4_10_0pareto : 0.046 seconds
+sim3_featspacelen_4_10_1pareto : 0.045 seconds
+sim3_featspacelen_4_25_0pareto : 0.056 seconds
+sim3_featspacelen_4_25_1pareto : 0.037 seconds
+sim3_featspacelen_4_50_0pareto : 0.028 seconds
+sim3_featspacelen_4_50_1pareto : 0.071 seconds
+sim3_featspacelen_16_5_0pareto : 0.338 seconds
+sim3_featspacelen_16_5_1pareto : 0.419 seconds
+sim3_featspacelen_16_10_0pareto : 0.536 seconds
+sim3_featspacelen_16_10_1pareto : 0.699 seconds
+sim3_featspacelen_16_25_0pareto : 0.621 seconds
+sim3_featspacelen_16_25_1pareto : 0.515 seconds
+sim3_featspacelen_16_50_0pareto : 0.584 seconds
+sim3_featspacelen_16_50_1pareto : 0.752 seconds
+sim3_featspacelen_32_5_0pareto : 2.014 seconds
+sim3_featspacelen_32_5_1pareto : 2.501 seconds
+sim3_featspacelen_32_10_0pareto : 2.142 seconds
+sim3_featspacelen_32_10_1pareto : 2.176 seconds
+sim3_featspacelen_32_25_0pareto : 2.277 seconds
+sim3_featspacelen_32_25_1pareto : 2.080 seconds
+sim3_featspacelen_32_50_0pareto : 2.555 seconds
+sim3_featspacelen_32_50_1pareto : 1.944 seconds
+sim3_featspacelen_64_5_0pareto : 8.940 seconds
+sim3_featspacelen_64_5_1pareto : 10.489 seconds
+sim3_featspacelen_64_10_0pareto : 9.777 seconds
+sim3_featspacelen_64_10_1pareto : 8.458 seconds
+sim3_featspacelen_64_25_0pareto : 10.321 seconds
+sim3_featspacelen_64_25_1pareto : 9.168 seconds
+sim3_featspacelen_64_50_0pareto : 9.095 seconds
+sim3_featspacelen_64_50_1pareto : 8.893 seconds
 

### 2013-01-03 ParetoHMM/sim3.py

@@ -64,6 +64,13 @@ class SimExp(object) :
 			fname = 'data/sim'+str(STUDY)+'_'+self.name+'_'+\
 				'_'+str(self.kwdargs['seqspace'])+\
 				'_'+str(self.kwdargs['seqlen'])+'.pkl'
+		if self.name == 'featspace' : 
+			fname = 'data/sim'+str(STUDY)+'_'+self.name+'_'+\
+				str(self.kwdargs['featspace'])+'.pkl'
+		elif  self.name == 'featspacelen' : 
+			fname = 'data/sim'+str(STUDY)+'_'+self.name+'_'+\
+				'_'+str(self.kwdargs['featspace'])+\
+				'_'+str(self.kwdargs['seqlen'])+'.pkl'	
 		else :
 			fname = 'data/sim'+str(STUDY)+'_'+self.name+'.pkl'
 
@@ -431,12 +438,69 @@ class SimExp(object) :
 			self.tasklist.append(task)	
 
 	def featspace(self) :	
-		""" Vary the feature space"""
-		pass
+		""" Vary sequence space and length"""
+		self.tasklist = []
+		featspace = self.kwdargs['featspace']
+		seqspace = 20
+		seqlen = 16
+		dims = [(seqspace,featspace)]*seqlen
 
+		# Repeat for all the tasks described	
+		for taskid in range(self.ntimes) :	
+			hmm = HMM()
+			self._set_params_generic(hmm,seqlen,dims)
+			cmrf = CMRF(hmm)	
+			feats = self._gen_feats_generic(seqlen,featspace)
+			task = Task('sim'+STUDY+'_'+self.name+'_'+\
+				str(featspace)+'_'+str(taskid),cmrf,feats)				
+			# Run Brute force to enumerate the frontier
+			if self.kwdargs['run_brute'] : 
+				with benchmark(task.name+'brute') as t:
+					seq,energies = self.bruteforce(cmrf,feats)			
+				task.all_seq = seq
+				task.all_seq_energy = energies
+				task.brute_time = t.elapsed			
+
+			# Now run the toy simulation`
+			with benchmark(task.name+'pareto') as t : 
+				task.frontier,task.frontier_energy = \
+					pareto_frontier(cmrf,feats)		
+			if self.plot_all :
+				task.plot_frontier(frontier_only=True)
+			task.pareto_time = t.elapsed
+			self.tasklist.append(task)	
  	def featspacelen(self) : 
[…]

### 2013-01-04 docs/tex/experiments.tex
Replaced with verbatim outputs

--- 2013-01-03 docs/tex/experiments.tex
@@
 \subsubsection{Simulation Experiment 5 : Random feature space tied}
-Like~\ref{sim:randprobstied} but the feature space is random
+Like~\ref{sim:randprobstied} but the feature space is randomized while the parameters between differnent amino acids are tied together. 
 \label{sim:randfeatstied}
@@
 \subsubsection{Simulation Experiment 6 :  Random feature space untied}
-Like~\ref{sim:randprobsuntied} but the feature space is random
+Like~\ref{sim:randprobsuntied} but the feature space is randomized while the different amino acids are untied. This is the most general of the previous settings. 
 \label{sim:randfeatsuntied}
+
+\begin{center}
+    \begin{tabular}{| c | c | c | c |}
+    \hline
+    Experiment & Brute & Pareto \\ \hline
+    Sim~\ref{sim:toy} & 0.907 & 0.003 \\ \hline
+    Sim~\ref{sim:randprobs} & 0.986 & 0.006 \\ \hline
+    Sim~\ref{sim:randprobstied} & 0.912 & 0.004 \\ \hline
+    Sim~\ref{sim:randprobsuntied} & 0.903 & 0.0034\\ \hline
+    Sim~\ref{sim:randfeatstied} & 0.921 & 0.011 \\ \hline
+    Sim~\ref{sim:randfeatsuntied} & 0.910 & 0.006 \\ \hline
+    \hline
+    \end{tabular}
+\end{center}
 
@@
 Increase the sequence space using random untied feature space as in ~\ref{sim:randfeatsuntied}. Will supply different seq space values. 2,4,16,20 are the values that are goint to be used.
+\begin{verbatim}
+sim3_seqspace_2_0pareto : 0.017 seconds
+sim3_seqspace_2_1pareto : 0.005 seconds
+sim3_seqspace_2_2pareto : 0.007 seconds
+sim3_seqspace_4_0pareto : 0.022 seconds
+sim3_seqspace_4_1pareto : 0.021 seconds
+sim3_seqspace_4_2pareto : 0.015 seconds
+sim3_seqspace_8_0pareto : 0.040 seconds
+sim3_seqspace_8_1pareto : 0.069 seconds
+sim3_seqspace_8_2pareto : 0.061 seconds
+sim3_seqspace_16_0pareto : 0.338 seconds
+sim3_seqspace_16_1pareto : 0.213 seconds
+sim3_seqspace_16_2pareto : 0.214 seconds
+sim3_seqspace_20_0pareto : 0.370 seconds
+sim3_seqspace_20_1pareto : 0.463 seconds
+sim3_seqspace_20_2pareto : 0.284 seconds
+\end{verbatim}
 
@@
 \label{sim:seqlen}
-Set the sequence space to be 20 and vary the sequence length from 2^2,2^16
+Set the sequence space to be 20 and vary the sequence length from $2^2$  to $2^7$
+\begin{verbatim}
+sim3_seqlen_4_0pareto : 0.047 seconds
+sim3_seqlen_4_1pareto : 0.011 seconds
+sim3_seqlen_4_2pareto : 0.029 seconds
+sim3_seqlen_16_0pareto : 0.621 seconds
+sim3_seqlen_16_1pareto : 0.857 seconds
+sim3_seqlen_16_2pareto : 0.499 seconds
+sim3_seqlen_32_0pareto : 2.343 seconds
+sim3_seqlen_32_1pareto : 2.344 seconds
+sim3_seqlen_32_2pareto : 2.409 seconds
+sim3_seqlen_64_0pareto : 8.683 seconds
+sim3_seqlen_64_1pareto : 10.716 seconds
+sim3_seqlen_64_2pareto : 11.068 seconds
+sim3_seqlen_128_0pareto : 39.451 seconds
+sim3_seqlen_128_1pareto : 40.050 seconds
+sim3_seqlen_128_2pareto : 33.411 seconds
+\end{verbatim}
+%
 
@@
 Jointly vary the sequence space and sequence length, 
+\begin{verbatim}
+sim3_seqspacelen_4_2_0pareto : 0.001 seconds
+sim3_seqspacelen_4_2_1pareto : 0.001 seconds
+sim3_seqspacelen_4_4_0pareto : 0.000 seconds
+sim3_seqspacelen_4_4_1pareto : 0.004 seconds
+sim3_seqspacelen_4_8_0pareto : 0.010 seconds
+sim3_seqspacelen_4_8_1pareto : 0.003 seconds
+sim3_seqspacelen_4_16_0pareto : 0.032 seconds
+sim3_seqspacelen_4_16_1pareto : 0.014 seconds
+sim3_seqspacelen_16_2_0pareto : 0.019 seconds
+sim3_seqspacelen_16_2_1pareto : 0.017 seconds
+sim3_seqspacelen_16_4_0pareto : 0.040 seconds
+sim3_seqspacelen_16_4_1pareto : 0.039 seconds
+sim3_seqspacelen_16_8_0pareto : 0.144 seconds
+sim3_seqspacelen_16_8_1pareto : 0.102 seconds
+sim3_seqspacelen_16_16_0pareto : 0.375 seconds
+sim3_seqspacelen_16_16_1pareto : 0.398 seconds
+sim3_seqspacelen_32_2_0pareto : 0.012 seconds
+sim3_seqspacelen_32_2_1pareto : 0.044 seconds
+sim3_seqspacelen_32_4_0pareto : 0.109 seconds
+sim3_seqspacelen_32_4_1pareto : 0.156 seconds
+sim3_seqspacelen_32_8_0pareto : 0.290 seconds
+sim3_seqspacelen_32_8_1pareto : 0.389 seconds
+sim3_seqspacelen_32_16_0pareto : 1.166 seconds
+sim3_seqspacelen_32_16_1pareto : 1.821 seconds
+sim3_seqspacelen_64_2_0pareto : 0.195 seconds
+sim3_seqspacelen_64_2_1pareto : 0.191 seconds
+sim3_seqspacelen_64_4_0pareto : 0.567 seconds
+sim3_seqspacelen_64_4_1pareto : 0.458 seconds
+sim3_seqspacelen_64_8_0pareto : 1.594 seconds
+sim3_seqspacelen_64_8_1pareto : 1.679 seconds
+sim3_seqspacelen_64_16_0pareto : 7.087 seconds
+sim3_seqspacelen_64_16_1pareto : 5.645 seconds
+sim3_seqspacelen_128_2_0pareto : 0.713 seconds
+sim3_seqspacelen_128_2_1pareto : 0.670 seconds
+sim3_seqspacelen_128_4_0pareto : 2.135 seconds
+sim3_seqspacelen_128_4_1pareto : 2.418 seconds
+sim3_seqspacelen_128_8_0pareto : 6.247 seconds
+sim3_seqspacelen_128_8_1pareto : 6.642 seconds
+sim3_seqspacelen_128_16_0pareto : 23.922 seconds
+sim3_seqspacelen_128_16_1pareto : 24.432 seconds
+\end{verbatim}
 
@@
 \label{sim:featspace}
-Keeping the seq space at 20, vary feat space from 5-10-25-50
+Keeping the seq space at 20 and seq len at 16. vary feat space from 5-10-25-50
+
+\begin{verbatim}
+sim3_featspace_5_0pareto : 0.344 seconds
+sim3_featspace_5_1pareto : 0.426 seconds
+sim3_featspace_5_2pareto : 0.499 seconds
+sim3_featspace_10_0pareto : 0.543 seconds
+sim3_featspace_10_1pareto : 0.706 seconds
+sim3_featspace_10_2pareto : 0.657 seconds
+sim3_featspace_25_0pareto : 0.632 seconds
+sim3_featspace_25_1pareto : 0.493 seconds
+sim3_featspace_25_2pareto : 0.536 seconds
+sim3_featspace_50_0pareto : 0.543 seconds
+sim3_featspace_50_1pareto : 0.676 seconds
+sim3_featspace_50_2pareto : 0.586 seconds
+\end{verbatim}
 
@@
 \label{sim:featspacelen}
-Keeping the seq space at 20, vary feat space from 5-10-25-50, sequence length from 2^2,2^16
+Keeping the seq space at 20, vary feat space from 5-10-25-50, sequence length from seqlen to 64.
 
+\begin{verbatim}
+sim3_featspacelen_4_5_0pareto : 0.030 seconds
+sim3_featspacelen_4_5_1pareto : 0.037 seconds
+sim3_featspacelen_4_10_0pareto : 0.046 seconds
+sim3_featspacelen_4_10_1pareto : 0.045 seconds
+sim3_featspacelen_4_25_0pareto : 0.056 seconds
+sim3_featspacelen_4_25_1pareto : 0.037 seconds
+sim3_featspacelen_4_50_0pareto : 0.028 seconds
+sim3_featspacelen_4_50_1pareto : 0.071 seconds
+sim3_featspacelen_16_5_0pareto : 0.338 seconds
+sim3_featspacelen_16_5_1pareto : 0.419 seconds
+sim3_featspacelen_16_10_0pareto : 0.536 seconds
+sim3_featspacelen_16_10_1pareto : 0.699 seconds
+sim3_featspacelen_16_25_0pareto : 0.621 seconds
+sim3_featspacelen_16_25_1pareto : 0.515 seconds
+sim3_featspacelen_16_50_0pareto : 0.584 seconds
+sim3_featspacelen_16_50_1pareto : 0.752 seconds
+sim3_featspacelen_32_5_0pareto : 2.014 seconds
+sim3_featspacelen_32_5_1pareto : 2.501 seconds
+sim3_featspacelen_32_10_0pareto : 2.142 seconds
+sim3_featspacelen_32_10_1pareto : 2.176 seconds
+sim3_featspacelen_32_25_0pareto : 2.277 seconds
+sim3_featspacelen_32_25_1pareto : 2.080 seconds
+sim3_featspacelen_32_50_0pareto : 2.555 seconds
+sim3_featspacelen_32_50_1pareto : 1.944 seconds
+sim3_featspacelen_64_5_0pareto : 8.940 seconds
+sim3_featspacelen_64_5_1pareto : 10.489 seconds
+sim3_featspacelen_64_10_0pareto : 9.777 seconds
+sim3_featspacelen_64_10_1pareto : 8.458 seconds
+sim3_featspacelen_64_25_0pareto : 10.321 seconds
+sim3_featspacelen_64_25_1pareto : 9.168 seconds
+sim3_featspacelen_64_50_0pareto : 9.095 seconds
+sim3_featspacelen_64_50_1pareto : 8.893 seconds
+\end{verbatim}
+%
 \subsubsection{Simulation Experiment : K-way multistate design}

### 2013-01-05 docs/tex/draft.tex
Simulation 7, seqspace
Added section on exploring concavities

@@
 
+\subsubsection{Side chain placement}
+We can use this for side chain placement by increasing the seq space of a particular amino acid to include rotameric energies.
+The node and edge potentials can encode energies for side chain placement.
+
+\subsection{Exploring Concavities }
+There might occur situations where the protein sequence lies in the concavities of the sequence space and is not strictly dominated.
+Strict domination means that the energies lie in the upper right block of all points of the points on the frontier.
+Figure~\ref{fig:cvxfront3} illustrates the point that there might be points in the interior of the convex frontier found by the algorithm that are not strictly dominated.
+P1 and P2 are on the convex frontier and the dashed magenta line is the convex hull.
+Any point in D1 is strictly dominated by P1 and any point in D2 is strictly dominated by P2.
+Therefore the points P3 and P4 are dominated by P1 and P2.
+The point P5 on the other hand is not dominated by P1 or P2 and yet is not explored by the algorithm.
+It does not lie on the convex frontier but it may lie on the pareto frontier.
+This is because P5 lies in the concavities of the sequence space.
+There is a way to explore the concavities of the sequence space if so desired.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[width=0.9\textwidth]{pics/convexfront3.pdf}
+    \caption{This figure illustrates the point that there might be points in the interior of the convex frontier found by the algorithm that are not strictly dominated.
+P1 and P2 are on the convex frontier and the dashed magenta line is the convex hull.
+Any point in D1 is strictly dominated by P1 and any point in D2 is strictly dominated by P2.
+Therefore the points P3 and P4 are dominated by P1 and P2.
+The point P5 on the other hand is not dominated by P1 or P2 and yet is not explored by the algorithm.
+It does not lie on the convex frontier but it may lie on the pareto frontier.
+This is because P5 lies in the concavities of the sequence space.}
+    \label{fig:cvxfront3}
+\end{figure}
+\pagebreak
+
 \section{Conclusion}

### 2013-01-05 docs/tex/experiments.tex
Added experiments biological
Added experiment descriptions

@@
+\documentclass{article}
+\usepackage{graphicx}
+\usepackage{amsmath,amssymb,amsthm}
+%\usepackage[square]{natbib}
+\usepackage{color}
+\usepackage{algpseudocode}
+\usepackage{algorithm}
+\usepackage{multirow}
+\usepackage{color}
+\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+
+\DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\argmin}{arg\,min}
+
+\begin{document}
+
 \section{Experiments}
@@
 
-\subsection{Simulation Experiment 11 : Augmenting Feature space}
+\subsubsection{Simulation Experiment 11 : Augmenting Feature space}
 
@@
 \subsection{Experiments on Biological Data}
+In this section we apply our algorithm to biological sequence data. These are the following main steps required :
+
+\begin{itemize}
+\item Create alignments of sequences 
+\item Train HMM from the alignment
+\item Convert HMM representation to CMRF
+\item Choose design state and corresponding features
+\item Run algorithm to decode sequence
+\end{itemize}
+
+There can be a number of different schemes that can be used for running biological experiments.
+
+\paragraph{Tied - SCOP} This scheme is the simplest of all.
+Try to find a number of proteins that belong to a particular protein fold.
+There should be sequence and structure information available for all these structures.
+Find the transition probabilities of these sequences and the corresponding sec.
+structure features.
+Use these to learn $P(X)$ and $P(X1|X2)$ and $P(O|X)$ that are tied for all amino acid positions.
+Find the 3D features of the states you are interested in designing.
+Could be the two states of Zif268 and Trimeric or Janus series or disordered proteins.
+
+\paragraph{Seq untied, features tied - PFAM/SCOP} -Similar to previous scheme but with sequence weights untied.
+We obtain PFAM alignments for the states in questions.
+Combine the alignments.
+Learn P(X) and P(X1|Xj).
+P(O|X) is learned from a general SCOP superfamily.
+Find 3D features of all positions you are interested in.
+Make sure that gaps and dummy features are handled carefully.
+Decode on a possibly larger HMM than the sequence desired because the MSA is bigger.
+
+\paragraph{Seq untied, features partially untied - PFAM/SCOP} - An effort is made to untie 3D features.
+Instead of having a uniform $P(O|X)$ for all sequence positions.
+Whichever sequences in alignment have 3D information, those sequences then bias $P(O|X)$ for that position.
+
+\paragraph{Seq, features untied - PDB Blast} - All sequence positions are untied. PDB is blasted to find all sequences that match given region of interest.
+
+\paragraph{No Joint alignment} - Have different alignments for the different states and learn separate HMMs for each of these. The algorithm needs to be modified in order to perform decoding on these.
+
+\paragraph{Augment Feature space discrete features} - Augment the feature space with other discrete features.
+
+\paragraph{Augment feature space real features }  - Augment feature space with real valued features. Condition on them using Bayes theorem and some a distribution such as the gaussian.
+
+\subsubsection{Tied SCOP - Disordered proteins}
+\label{exp:tied_scop_disordered}
+The goal is to be able to find where known disordered proteins lie on the sequence space defined by the family of disordered proteins.
+To do this we first obtain sequences corresponding to the family of disordered   proteins probably from SCOP.
+The sequences are examined to learn transition probabilities and begin probabilities.
+The pdb structures are examined to find the amino acid to secondary structure counts.
+The Emission and transition tables are filled in.
+
+There is no concept of gaps here.
+It is possible that different sequence lengths correspond to different energy values, sop the scores must be normalized to the length of the target design sequences.
+The decoding is performed and sequences are examined with respect to the pareto-frontier.
+
+\subsubsection{Untied SCOP Partial - Disordered proteins}
+\label{exp:untied_scop_disordered}
+The goal here is the same as in section~\ref{exp:tied_scop_disordered}.
+However we will do away with the concept of tied position independent hmms.
+Instead we will create a giant alignment of all PFAM/SCOP sequences corresponding to these fellows.
+We will use background frequencies of amino acids and features and use them as pseudo-counts.
+Those sequences that have pdb structures will be used to bias the emission probability weights where suitable.
+There will be a lot of aligning and re-aligning going on.
+So it will help to make sure that all this is taken care off properly.
+
+\subsubsection{Tied SCOP - Multistate proteins}
+
+\subsubsection{Untied SCOP partial- Multistate proteins}
+
+\subsection{Untied PDB Blast - Multistate proteins}
+
+\bibliographystyle{plain}
+\bibliography{refs}
+
+\end{document}

### 2013-01-06 docs/tex/experiments.tex
Added utils pdb fetch command and redefined experiments

@@
 
+We have the following sources of information from the various databases :
+
+\begin{itemize}
+\item pdb - pfam map
+\item pdb - scop - pdb list map
+\item pdb - sec structure map
+\item pdb - seq map
+\item pdb - pdb structure 
+\item pfam - pdb 
+\item pfam - uniprot (alignments)
+\end{itemize}
+
 There can be a number of different schemes that can be used for running biological experiments.
@@
 
+\subsubsection{Tied SCOP Zif268-Trimer}
+\label{exp:tied_scop_multistate}
+For multistate design we will look try to design for two states Zif268(zic-finger like) and a trimeric coiled coil.
+Specifically we will consider the structures of hemaglutinin(13-44) pdb id 1HTM and Zif268 zinc-finger DNA complex(pdb id: 1AAY).
+1HTM belongs to the PF00509 family which has 232 structures.
+1AAY belongs to PF00096 family which has 19 structures.
+These sets are expanded by considering the SCOP family that they belong to.
+1AAY belongs to the Zif268 domain which has 14 structures and the zinc finger family which has 70 structures.
+1HTM belongs to influenza hemagluttinin family and domain which have 39 structures both.
+In order to train the hmm, one can learn the transition and emission probabilities of the region under consideration, the whole protein.
+One can also choose to learn these weights from just the PFAM or just the scop data or both combined.
+We will explore each of these cases.
+
+\subsubsection{Untied SCOP-PFAM Zif268-Trimer}
+\label{exp:untied_scop_zif}
+
 \subsubsection{Tied SCOP - Disordered proteins}
@@
 
-\subsubsection{Tied SCOP - Multistate proteins}
-
 \subsubsection{Untied SCOP partial- Multistate proteins}

### 2013-01-06 ParetoHMM/utils.py

@@ -1,0 +1,12 @@
+""" 
+A bunch of utility scripts
+"""
+
+import urllib2
+
+def fetch_pdb(id):
+  url = 'http://www.rcsb.org/pdb/files/%s.pdb' % id
+  return urllib2.urlopen(url).read()
+
+
+

### 2013-01-06 ParetoHMM/exp1.py
Added skeleton for exp1

@@ -1,0 +1,75 @@
+""" 
+In this experiment we will be enumerating the pareto frontier for 
+the Zif268 multistate design problem
+
+"""
+
+import os,sys
+from pdb import set_trace as stop
+import numpy as np
+import pylab as pl
+import scipy as sci
+import subprocess
+
+# Constants
+pdb_ss_fpath = 'data/ss_dis.txt'
+SEEK_BUFF = 100
+
+
+def build_tied_hmm(self,align_fpath,feats) :
+	""" Build a HMM using data from an alignment """
+
+def extract_feats(pdbid) : 
+	""" Extract seq and ss from the filename"""
+	global pdb_ss_fpath
+	pdbid = pdbid.upper()
+	out = subprocess.check_output(['grep','-n',pdbid,pdb_ss_fpath])
+	skip = out.split()[0].split(':')[0]
+	p1 = subprocess.Popen(['head','-n','+'+str(int(skip)-1+SEEK_BUFF),\
+		pdb_ss_fpath],stdout=subprocess.PIPE)
+	p2 = subprocess.Popen(['tail','-n',str(SEEK_BUFF)],stdin=p1.stdout,\
+		stdout=subprocess.PIPE)
+	p1.stdout.close()
+	output = p2.communicate()[0]
+	
+	# Extract the sequence and the ss
+	state=0
+	seq,ss,dis = '','',''
+	for line in output.split('\n'):
+		if state==0  and line[0] == '>' and 'sequence' in line : 
+			state = 1
+			continue
+		if state==1 :
+			if line[0] == '>' and 'secstr' in line :
+				state = 2
+				continue 
+			else : 
+				seq += line.strip('\n')
+				continue	
+		if state==2 :
+			if line[0] == '>' and 'disorder' in line :
+				state = 3
+				continue 
+			else : 
+				ss += line.strip('\n')
+				continue	
+		if state == 3 : 
+			if line[0] == '>' : 
+				break
+			else : 
+				dis += line.strip('\n')
+	ss = ss.replace(' ','L')
+	return seq,ss,dis	
+
+class Protein(object) :
+	""" Describes a protein class """
+	def __init__(self,pdbid,design_range) : 
+		self.pdbid = pdbid
+		self.seq,self.ss,self.dis = extract_feats(pdbid)
+		self.design_range = design_range	
+		self.des_feat = self.ss[design_range[0][0]-1:design_range[0][1]]
[…]

### 2013-01-06 ParetoHMM/exp1.py
Wrote scop file parser

@@ -10,9 +10,14 @@ import numpy as np
 import pylab as pl
 import scipy as sci
 import subprocess
+import re
 
 # Constants
 pdb_ss_fpath = 'data/ss_dis.txt'
+pdb_seq_fpath = 'data/pdb_seqres.txt'
+pdb_scop_1htm_fpath = 'data/1htm_scop.list'
+pdb_scop_1aaw_fpath = 'data/1aaw_dom_scop.list'
+
 SEEK_BUFF = 100
 
 
@@ -61,15 +66,58 @@ def extract_feats(pdbid) :
 	ss = ss.replace(' ','L')
 	return seq,ss,dis	
 
+def extract_pdb_seq(pdbid) : 
+	pass		
+
 class Protein(object) :
 	""" Describes a protein class """
 	def __init__(self,pdbid,design_range) : 
 		self.pdbid = pdbid
 		self.seq,self.ss,self.dis = extract_feats(pdbid)
-		self.design_range = design_range	
-		self.des_feat = self.ss[design_range[0][0]-1:design_range[0][1]]
-		self.des_seq = self.seq[design_range[0][0]-1:design_range[0][1]]
+		if design_range == 'all' : 
+			self.design_range = [(1,len(self.seq))]
+		else : 
+			self.design_range = design_range	
+		self.des_feat = self.ss[self.design_range[0][0]-1:\
+			self.design_range[0][1]]
+		self.des_seq = self.seq[self.design_range[0][0]-1:\
+			self.design_range[0][1]]
+
+def read_pdb_scop_list(fpath) : 
+	""" Read a pdb list file and return a list of pdb ids """
+	with open(fpath) as fin : 
+		pdb_list = [line.strip() for line in fin.readlines()]
+	return pdb_list
+
+def parse_scop_csv(fpath) :
+	""" Parse the scop file and return the data """
+	parsed = {'pdbid':[],'chain':[],'pfam':[],'seq':[],'sec':[]}
+	regex = re.compile(r'\"(.*?)\"')
+
+	with open(fpath) as fin : 
+		headers = fin.readline() # skip the first line
+		for line in fin.readlines() :
+			if line == '\n' :
+				continue
+			line_list = regex.findall(line.strip('\n'))
+			pdbid,chain,pfam,seqsec = line_list
+			seq,sec=seqsec.split('#')
+			sec = sec.replace(' ','L')
[…]

### 2013-01-06 ParetoHMM/exp1.py
Completely prepare scop dataset

@@ -11,6 +11,7 @@ import pylab as pl
 import scipy as sci
 import subprocess
 import re
+from operator import itemgetter
 
 # Constants
 pdb_ss_fpath = 'data/ss_dis.txt'
@@ -118,6 +119,18 @@ if __name__ == '__main__' :
 	scop_1htm = parse_scop_csv('data/1htm_scop.csv')
 	scop_1aay = parse_scop_csv('data/1aay_scop.csv')	
 
+ 	# Remove duplicate sequence and structures and store that \
+	#as the alignment
+	nondup_1htm = set(zip(scop_1htm['seq'],scop_1htm['sec']))
+	nondup_1aay = set(zip(scop_1htm['seq'],scop_1htm['sec']))
 
-	
-	
+	# Get sequences and ss for each of the
+	data = {}
+	data['1htm'] = {}
+	data['1aay'] = {}
+	data['1htm']['seq'] = map(itemgetter(0),nondup_1htm)
+	data['1aay']['seq'] = map(itemgetter(0),nondup_1aay)
+	data['1htm']['ss'] = map(itemgetter(1),nondup_1htm)
+	data['1aay']['ss'] = map(itemgetter(1),nondup_1aay)
+
+

### 2013-01-06 ParetoHMM/exp1.py
Wrote method for calculating base frequencies

@@ -67,9 +67,6 @@ def extract_feats(pdbid) :
 	ss = ss.replace(' ','L')
 	return seq,ss,dis	
 
-def extract_pdb_seq(pdbid) : 
-	pass		
-
 class Protein(object) :
 	""" Describes a protein class """
 	def __init__(self,pdbid,design_range) : 
@@ -109,6 +106,30 @@ def parse_scop_csv(fpath) :
 				eval('parsed["%s"].append(%s)'%(key,key))	
 	return parsed
 
+
+def calc_base(data)  :
+	""" Calculate all the base frequencies of all chars in dataset"""
+	# Scop classes
+	scop_classes = data.keys()
+
+	counts = {}
+	# Count the freqs
+	for c in scop_classes : 
+		cseq = data[c]['seq']
+		css = data[c]['ss']
+		counts_seq = {}
+		counts_ss = {}
+		for seq in cseq : 
+			for aa in seq : 
+				counts_seq[aa] = counts_seq.get(aa,0)+1
+		for seq in css : 
+			for aa in seq : 
+				counts_ss[aa] = counts_ss.get(aa,0)+1
+		counts[c] = {}
+		counts[c]['seq'] = counts_seq
+		counts[c]['ss'] = counts_ss
+
+	return counts
 	
 if __name__ == '__main__' : 
 	# Create features
@@ -133,4 +154,7 @@ if __name__ == '__main__' :
 	data['1htm']['ss'] = map(itemgetter(1),nondup_1htm)
 	data['1aay']['ss'] = map(itemgetter(1),nondup_1aay)
 
+	# Learn the weights of the hmm
+	base_data = calc_base(data)
+	
 

### 2013-01-06 docs/tex/experiments.tex
Wrote method for calculating base frequencies
Designed tied experiment for zif268. Experiment failed with slope>0. Can't imagine why

@@
 
+We use all the sequences and secondary structures of SCOP fold of the protein under consideration.
+For 1HTM it is the 'Stalk segment of viral fusion proteins' fold.
+For 1AAY it is 'beta-beta-alpha zinc fingers' fold.
+While training the HMM we deduplicate all sequence, sec.
+structure pairs.
+All X and U amino acids are not considered.
+We add a 5\% bias to all background to match background frequencies.
+This is done by :
+\[
+P(O=a|X=b) = \frac{Count(O=a,X=b)+Count(O=a)\times 0.05}{Count(X=b)+Count(O)\times 0.05}
+\]
+
 \subsubsection{Untied SCOP partial- Multistate proteins}

### 2013-01-06 ParetoHMM/exp1.py

@@ -12,6 +12,10 @@ import scipy as sci
 import subprocess
 import re
 from operator import itemgetter
+from itertools import product
+from HMM import HMM,CMRF, TMRF, reverse_dict
+import random,time, pickle
+from cvxhull import pareto_frontier
 
 # Constants
 pdb_ss_fpath = 'data/ss_dis.txt'
@@ -20,7 +24,12 @@ pdb_scop_1htm_fpath = 'data/1htm_scop.list'
 pdb_scop_1aaw_fpath = 'data/1aaw_dom_scop.list'
 
 SEEK_BUFF = 100
+DSSP_SYM = ('H','B','E','G','I','T','S','L')
+SEQ_SYM='ACDEFGHIKLMNPQRSTVWY'
+BIAS=0.05
+IGNORE_AA = 'XU'
 
+STUDY=1
 
 def build_tied_hmm(self,align_fpath,feats) :
 	""" Build a HMM using data from an alignment """
@@ -113,27 +122,267 @@ def calc_base(data)  :
 	scop_classes = data.keys()
 
 	counts = {}
+	counts['all'] = {}
+	counts['all']['seq'] = {}
+	counts['all']['ss'] = {}
 	# Count the freqs
 	for c in scop_classes : 
 		cseq = data[c]['seq']
 		css = data[c]['ss']
 		counts_seq = {}
 		counts_ss = {}
-		for seq in cseq : 
-			for aa in seq : 
+		for seq,ss in zip(cseq,css) : 
+			for aa,ss_aa in zip(seq,ss) : 
+				if aa in IGNORE_AA : 
+					continue
 				counts_seq[aa] = counts_seq.get(aa,0)+1
-		for seq in css : 
-			for aa in seq : 
-				counts_ss[aa] = counts_ss.get(aa,0)+1
+				counts_ss[ss_aa] = counts_ss.get(aa,0)+1
 		counts[c] = {}
 		counts[c]['seq'] = counts_seq
 		counts[c]['ss'] = counts_ss
+		for k in counts[c]['seq'].keys() : 
+			counts['all']['seq'][k] =  counts['all']['seq'].get(k,0) +\
+				counts[c]['seq'][k]	
+
+		for k in counts[c]['ss'].keys() :
+			counts['all']['ss'][k] =  counts['all']['ss'].get(k,0) +\
+				counts[c]['ss'][k]	
 
 	return counts
 	
+def learn_weights_combined(data,base_data) :
[…]

### 2013-01-07 ParetoHMM/cvxhull.py
Fixed the bug that was causing m>0 problems. Basically there are some points which are very close to other points and that causes the floating point arithmetic and other stuff to get confused

@@ -11,6 +11,10 @@ from HMM import HMM,CMRF, TMRF
 from pdb import set_trace as stop
 import warnings
 
+def almost_eq(e1,e2) : 
+	""" Check if two elements are almost equal """
+	return round(e1-e2,4) == 0.0
+
 def pareto_frontier(cmrf,featlist) :
 	"""Finds and prints the pareto frontier. Currently works for 
 		two competing states only
@@ -51,7 +55,12 @@ def pareto_frontier(cmrf,featlist) :
 		Eaxab = cmrf.score(Xab,feat1)
 		Ebxab = cmrf.score(Xab,feat2)
 		if Xab != Xa and Xab != Xb and \
-			not taboodict.has_key((Eaxab,Ebxab)): 
+			not taboodict.has_key((Eaxab,Ebxab)) :
+			# Check almost equal condition
+			if any(map(lambda(x):almost_eq(Eaxab,x[0] or \
+				almost_eq(Ebxab,x[1])),taboodict.keys())) : 
+				continue
+					
 			frontier.append(Xab)
 			frontier_energy.append((Eaxab,Ebxab))
 			taboodict[(Eaxab,Ebxab)]=1

### 2013-01-07 ParetoHMM/exp1.py

@@ -28,7 +28,7 @@ DSSP_SYM = ('H','B','E','G','I','T','S','L')
 SEQ_SYM='ACDEFGHIKLMNPQRSTVWY'
 BIAS=0.05
 IGNORE_AA = 'XU'
-
+SAMPLE_N = 10000
 STUDY=1
 
 def build_tied_hmm(self,align_fpath,feats) :
@@ -259,6 +259,7 @@ class BioExp(object) :
 		weights = self.kwdargs['weights']
 		hmm = HMM()
 		self._set_params_ziftied(hmm)
+		#1/0
 		cmrf = CMRF(hmm)
 		for taskid in range(self.ntimes) :	
 			task = Task('bio'+str(STUDY)+'_'+self.name+'_'+str(taskid),cmrf,\
@@ -270,12 +271,19 @@ class BioExp(object) :
 #			task.all_seq_energy = energies
 #			task.brute_time = t.elapsed			
 
+			# Sample the frontier
+			with benchmark(task.name+'sample') as t:
+				seq,energies = self.sample(cmrf,feats)			
+			task.sample_seq = seq
+			task.sample_seq_energy = energies
+			task.sample_time = t.elapsed			
+
 			# Now run the toy simulation`
 			with benchmark(task.name+'pareto') as t : 
 				task.frontier,task.frontier_energy = \
 					pareto_frontier(cmrf,feats)		
 			if self.plot_all :
-				task.plot_frontier(frontier_only = True)
+				task.plot_frontier(frontier_only = True,plot_samples=True)
 			task.pareto_time = t.elapsed
 			self.tasklist.append(task)	
 	
@@ -320,7 +328,23 @@ class BioExp(object) :
 			ll_list2.append(cmrf.score(seq,feat2))
 
 		return seq_list,zip(ll_list1,ll_list2)
-	
+
+	def sample(self,cmrf,feats)  :
+		""" Randomly sample a section of the sequence space """
+		feat1,feat2 = feats
+		ll_list1,ll_list2 = [],[]
+		seqdim = cmrf.dims[0][0]	
+		seqlen = cmrf.length
+		seqspace ='ACDEFGHIKLMNPQRSTVWY'[:seqdim]
+		for i in range(SAMPLE_N):
+			seq = "".join([random.choice(seqspace) for i in range(seqlen)])
+			ll_list1.append(cmrf.score(seq,feat1))
[…]

### 2013-02-12 docs/tex/draft.tex
Boosted simulations

@@
 
-\include{experiments}
+%\include{experiments}
 
@@
 This is because P5 lies in the concavities of the sequence space.
-There is a way to explore the concavities of the sequence space if so desired.
+
+In general, the convex pareto-frontier contains high quality sequences in itself.
+The convex hull contains sequences that are solutions for all values of $\theta$ for the standard linear multistate design problem $\argmin_X \theta E_A(X)+(1-\theta)E_B(X)$.
+However, there do lie undominated points in the concavities that maybe worthwhile to explore.
+Fortunately, there exists a way to explore the concavities as shown previously in a different context by Zheng et.al~\cite{Zheng2009}.
+The authors apply the convex hull procedure to local sections of the frontier by freezing entities in their dynamic programming problem.
+We modify their approach in the context of dynamic programming for TMRFs.
+Basically, the idea is to apply the Pareto-frontier method while keeping a residue to fixed to a certain amino acid.
+Since max-decoding can still be performed while keeping a node fixed this procedure is feasible.
+We then iterate over all amino acids and over all residues while performing max-decoding while culling dominated points.
+This will significantly explore the concavities.
+Keeping two or more residues fixed will further explore the space but an additional cost.
 

### 2013-02-12 ParetoHMM/exp1.py

@@ -413,7 +413,7 @@ class benchmark(object):
 
 if __name__ == '__main__' : 
 	# Create features
-	p1 = Protein('1HTM',[(13,33)])	
+	p1 = Protein('1HTM',[(13,43)])	
 	p2 = Protein('1AAY',[(3,33)])
 
 	# Read 1htm scop file

### 2013-03-07 ParetoHMM/exp2.py
sim4 for oversampling,,exp2.py for real data

@@ -1,0 +1,7 @@
+""" 
+In this section we do tied 2 -state design
+"""
+
+
+
+

### 2013-03-07 ParetoHMM/exp2.py
Experiment on real data, code for generating alignments

@@ -1,0 +1,646 @@
+""" 
+In this experiment we will be enumerating the pareto frontier for 
+the Zif268 multistate design problem
+
+"""
+
+import os,sys
+from pdb import set_trace as stop
+import numpy as np
+import pylab as pl
+import scipy as sci
+import subprocess
+import re
+from operator import itemgetter
+from itertools import product
+from HMM import HMM,CMRF, TMRF, reverse_dict
+import random,time, pickle
+from cvxhull import pareto_frontier
+from Bio import SeqIO,AlignIO
+from Bio.Seq import Seq
+from Bio.SeqRecord import SeqRecord 
+from Bio.Alphabet import generic_protein
+import copy
+
+
+# Constants
+pdb_ss_fpath = 'data/ss_dis.txt'
+pdb_seq_fpath = 'data/pdb_seqres.txt'
+pdb_scop_1htm_fpath = 'data/1htm_scop.list'
+pdb_scop_1aaw_fpath = 'data/1aaw_dom_scop.list'
+
+SEEK_BUFF = 100
+DSSP_SYM = ('H','B','E','G','I','T','S','L','D')
+SEQ_SYM='ACDEFGHIKLMNPQRSTVWY-'
+BIAS=0.05
+IGNORE_AA = 'XUB'
+SAMPLE_N = 10000
+STUDY=1
+ROSETTA_SEQS = [
+'RPYACPVESCDRRFSRSDRLTRHIRIHLRQK',
+'GKLNRVIEKTNEKFHQIEKEFSEVEGRIQDL',
+'KEFECLVEECEKKFKRLEELFRHVKEHIEKL',
+'KEFECEKEECEKKFKRLEEKLRHLEKHIKEL',
+'QEFECLKEECEQKFNQNNELDEHIRGHIQEL'
+]
+
+
+
+def build_tied_hmm(self,align_fpath,feats) :
+	""" Build a HMM using data from an alignment """
+
+def extract_feats(pdbid) : 
+	""" Extract seq and ss from the filename"""
+	global pdb_ss_fpath
+	pdbid = pdbid.upper()
+	out = subprocess.check_output(['grep','-n',pdbid,pdb_ss_fpath])
+	skip = out.split()[0].split(':')[0]
+	p1 = subprocess.Popen(['head','-n','+'+str(int(skip)-1+SEEK_BUFF),\
+		pdb_ss_fpath],stdout=subprocess.PIPE)
[…]

### 2013-03-07 ParetoHMM/gen_align.py

@@ -1,0 +1,62 @@
+"""
+Generate the aligns 
+"""
+
+from Bio import SeqIO,AlignIO
+from Bio.Seq import Seq
+from Bio.SeqRecord import SeqRecord
+from Bio.Alphabet import generic_protein
+
+
+map1htm = 'data/PF00509_full_1htm_map.fasta'
+full1htm = 'data/PF00509_full.fasta'
+full1htm_filt = 'data/full_1htm.fasta'
+map1aay = 'data/1aay_full.map'
+full1aay = 'data/PF00096_full.fasta'
+full1aay_filt = 'data/full_1aay.fasta'
+
+GAP_LIM=0.3 # Keep only sequences with less than 30% gap
+
+def gen_filter(pos,fpath) : 
+	for seqr in SeqIO.parse(fpath,"fasta") : 
+		seq = Seq("".join([seqr.seq[p] for p in pos]),generic_protein)
+		if (seq.count('-')+0.0)/len(seq) > GAP_LIM : 
+			continue
+		seqr2 = SeqRecord(seq,id=seqr.id)
+		yield seqr2
+
+def gen_filter2(pos,fpath) : 
+	for seqr in SeqIO.parse(fpath,"fasta") : 
+		seq = ""
+		for p in pos :
+			if p == '-' :
+				seq += '-'
+			else :
+				seq += seqr.seq[int(p)]
+		seq = Seq(seq,generic_protein)
+		if (seq.count('-')+0.0)/len(seq) > GAP_LIM : 
+			continue
+		seqr2 = SeqRecord(seq,id=seqr.id)
+		yield seqr2
+
+if __name__ == '__main__' : 
+	
+	#### Create the 1htm alignment
+	# Read the map
+#	seqs = [seqr.seq for seqr in SeqIO.parse(map1htm,"fasta")]
+#	seq = seqs[1]
+#	pos = [i for i,aa in enumerate(seq) if aa != '-']
+#
+#	# Load the large alignment
+#	gen_seqs = gen_filter(pos,full1htm)
+#	SeqIO.write(gen_seqs,full1htm_filt,"fasta")
+	
+
+	#### Create the 1aay alignment
+	# Read the map
+	with open(map1aay) as fin : 
+		pos = [line.strip().split()[0] for line in fin.readlines()]
+
+	# Load the large alignment
+	gen_seqs =  gen_filter2(pos,full1aay) 
+	SeqIO.write(gen_seqs,full1aay_filt,"fasta")
